### 2017-01-28 eswc-2017-review-response.md
Add ESWC 2017 review response

Dear Sarven,

Thank you for your submission to ESWC 2017. The ESWC 2017
review response period will be between now and Sunday 5th February 2017, midnight Hawaii time.

During this time, you will have access to the current state of
your reviews and have the opportunity to submit a response of
up to 1000 words. Please keep in mind the following during this
process:

* The response must focus on any factual errors in the reviews
and any questions posed by the reviewers. It must not provide
new research results or reformulate the presentation. Try to be
as concise and to the point as possible.

* The review response period is an opportunity to react to the
reviews, but not a requirement to do so. Thus, if you feel the
reviews are accurate and the reviewers have not asked any
questions, then you do not have to respond.

* The reviews are as submitted by the PC members, without any
coordination between them. Thus, there may be inconsistencies.
Furthermore, these are not the final versions of the reviews.
The reviews can later be updated to take into account the
discussions at the program committee meeting, and we may find
it necessary to solicit other outside reviews after the review
response period.

* The program committee will read your responses carefully and
take this information into account during the discussions. On
the other hand, the program committee will not directly respond
to your responses, either before the program committee meeting
or in the final versions of the reviews.

* Your response will be seen by all PC members who have access
to the discussion of your paper, so please try to be polite and
constructive.

The reviews on your paper are attached to this letter. To
submit your response you should log on the EasyChair Web page
for ESWC 2017 and select your submission on the menu.

----------------------- REVIEW 1 ---------------------
PAPER: 29
TITLE: Linked Data Notifications
AUTHORS: Sarven Capadisli, Amy Guy, Christoph Lange, Sören Auer and Tim Berners-Lee

Overall evaluation: 2

----------- Overall evaluation -----------
The article presents the Linked Data Notifications (LDN) protocol, a candidate W3C recommendation, and how it relates to similar notification protocols. Starting with requirements based on Linked Data design principles and the need for re-decentralisation of the Web the article makes a structured comparison of notification mechanisms today and discusses the various implementations of LDN.

This is an architecture type of article in which the requirements and design considerations (section 3) are the base on which the rest of the contribution is built; naturally, some of those requirements are based on opinion but this is exactly the kind of topic to present at ESWC to foster fruitful discussion. For example, the requirement for reusability could be seen as a potential risk for privacy; what is the benefit of reusable notifications to balance that risk? Who would benefit the most from this principle (users, application developers, other) and in what way? A similar discussion could be made for R3 (persistence and retrievability).

It would help if those requirements were discussed a little more in the paper and if they could be placed more clearly in the wider framework/paradigm of Web re-decentralisation (including stakeholders) to give a sense of ‘completeness’ when it comes to notification protocols.

One final point is that certain references (e.g. [3] do not seem to be cited in the text).

----------------------- REVIEW 2 ---------------------
PAPER: 29
TITLE: Linked Data Notifications
AUTHORS: Sarven Capadisli, Amy Guy, Christoph Lange, Sören Auer and Tim Berners-Lee

Overall evaluation: 3

----------- Overall evaluation -----------
LDNs provide an important mechanism for the current effort (as described in the paper) to re-decentralise the web. The paper is well written. It describes the motivation for LDNs, gives examples and compares LDNs to related efforts.

I have only minor comments:

1) one may add in section 2 "related work" a mention of atomPub https://tools.ietf.org/html/draft-snell-atompub-notification-01 , and add this also to table 2 "comparison of notification mechanisms"

2) for the standardisation process of LDNs this may not be crucial, but one may want to examine benchmarks for notification processing. This could help to evaluate different LDNs applications in terms of their performance. This criterion may become critical during broad adoption,  with billions of notifications generated every day.

----------------------- REVIEW 3 ---------------------
PAPER: 29
TITLE: Linked Data Notifications
AUTHORS: Sarven Capadisli, Amy Guy, Christoph Lange, Sören Auer and Tim Berners-Lee

Overall evaluation: 2

----------- Overall evaluation -----------
This paper presents a communication protocol for Linked Data Notifications, reusable messages on the Web that follow the Linked Data principles.

The paper is generally well written and easy to follow. The authors motivate their work well and provide a detailed description of the protocol and its implementations.

Strong points:
==============
- The requirements for the protocol are well established.
- The protocol itself is well formalised and described.
- Existence of a number of implementations, both standalone ones and the ones developed previously to the LDN.
- Protocol flexibility and the compatibility with the Linked Data Platform.
- The work is quite relevant.

Weak points:
============
- The lack of a real evaluation. Section 6, although titled "Analysis and Evaluation", is actually a comprehensive analysis and discussion of various aspects of the presented work. Section 6.1 is the closest to the evaluation of the work, however the level of details in this section is not satisfactory in that regard. The authors could maybe extend this section with an in depth analysis of the differences between their work and other system that are mentioned, and a discussion on why these differences are important and how LDN provides improvements over them.

Other comments:
==============

In the abstract, the authors state: "This permits end users more freedom to switch between the online tools they use, as well as generating greater value when notifications from different sources can be used in combination." Maybe it would be interesting if the authors could mention a few use cases of this added value in a separate section as a strong point for the motivation of the work.

In my opinion, the definition of notification is quite broad and not very clear ("A notification is a retrievable resource which returns RDF"). Is my URI a notification? A URI of an ontology class?
Further explanation perhaps clarifies a bit, but the whole context should be clear in the first sentence.

The interaction steps between consumer and receiver (Section 4.2) might be a bit confusing to the reader. Does step (3) always follow step (2), e.g., the consumer first asks for the listing of notifications and then choses one to retrieve, or it is possible to have only one of the two steps before step (4), e.g., the consumer decides whether it is interested in the listing, or individual notification, or perhaps both.
Also, in my opinion comparing to Section 4.1, the steps in Section 4.2 are described with less detail and rigour.

In Section 5 the authors state: "We note that any LDP implementation is a conforming LDN receiver; we refer here to the ones we have tested". Maybe the authors could mention the reasons why other LDP implementations are not tested, or how they have chosen the implementations that are tested.

The figures in Section 5 are not very readable in the printed version. I understand that the online version of the article reads better, but the authors should maybe consider improving the representations in printed form.

The conclusions section presents only a summary of the paper. But what are the conclusions? Are there any limitations of the presented work and if so which ones? Is there any future work planned?

Not all figures and tables are referenced in the text, I would suggest the authors to include these references.

Reference [3] is not referenced in the text.

In my opinion, the references section could be improved. For example, conference and workshop abbreviations could be extended, exact workshops mentioned (e.g., in [9]). Furthermore, links are omitted in the printed version of the paper which makes the section strange and it is not clear what is the type of the work that is referenced (e.g., conference, technical report, web article, technical specification)

----------------------- REVIEW 4 ---------------------
PAPER: 29
TITLE: Linked Data Notifications
AUTHORS: Sarven Capadisli, Amy Guy, Christoph Lange, Sören Auer and Tim Berners-Lee

Overall evaluation: 1

----------- Overall evaluation -----------
The paper is well written  and mostly understandable, though it might be a good idea to rephrase some of the explanation in the section on Requirements and Design Considerations, specifically R4-B, which is confusing and not easy to understand.

I would also like more clarity on the following:

1. Performance metrics to measure how well the protocol performs in practice.
2. The most interesting part of the protocol is the (re)use of the received notification by the consumers. The paper should include explicit justification with concrete examples on why an application (consumer) would want to use an outdated notification exposed by some receiver. Most obvious examples are blog posts etc, but that seems to be the only obvious use case. It is also worth thinking about how a service like stack overflow can be re-purposed using LDN and what would be the benefits.

It is also easy to see why a formal semantics for the protocol would make sense. Apart from getting implementations to conform to a test suite, it is worth getting them to conform to a formal semantics specification. This is something that the authors should 
make a part of their future work.

I also see the evaluation as being quite limited, as there is no user-driven evaluation result that have been reported. While it is still early days for the protocol, without a clear idea of what it means for the consumers and how much it contributes to improving the overall data integration for a specific use-case or domain, it is hard to see the utility of implementing this as part of a larger system.

------------------------------------------------------

Many Thanks!
The ESWC 2017 PC Chairs and Track Chairs.

### 2017-03-07 index.tex
Initial conversion

  \begin{document}

                Linked Data Notifications: a resource-centric communication protocol

                    \begin{description}
  \item[Authors] \empty Sarven Capadisli\empty 1\empty ✊\empty Amy Guy\empty 2\empty 🐦\empty Christoph Lange\empty 1\empty ∫\empty Sören Auer\empty 1\empty ⚛\empty Tim Berners-Lee\empty 3\empty 🕸
    \end{description}

                    \begin{itemize}
  \item 1Enterprise Information Systems Department, \empty University of Bonn, Bonn, DE\item 2School of Informatics, \empty University of Edinburgh, Edinburgh, UK\item 3Decentralized Information Group, CSAIL, \empty MIT, Cambridge, US
    \end{itemize}

                    \begin{itemize}
  \item ✊\empty info@csarven.ca\item 🐦\empty amy@rhiaro.co.uk\item ∫\empty langec@cs.uni-bonn.de\item ⚛\empty auer@cs.uni-bonn.de\item 🕸\empty timbl@w3.org
    \end{itemize}

                \begin{description}
  \item[Identifier] \empty http://csarven.ca/linked-data-notifications
    \end{description}

\par In this article we describe the Linked Data Notifications (LDN) protocol, which is a \empty W3C Candidate Recommendation. Notifications are sent over the Web for a variety of purposes, for example, by social applications. The information contained within a notification is structured arbitrarily, and typically only usable by the application which generated it in the first place. In the spirit of Linked Data, we propose that notifications should be reusable by multiple authorised applications. Through separating the concepts of {\em senders}, {\em receivers} and {\em consumers} of notifications, and leveraging Linked Data principles of shared vocabularies and URIs, LDN provides a building block for decentralised Web applications. This permits end users more freedom to switch between the online tools they use, as well as generating greater value when notifications from different sources can be used in combination. We situate LDN alongside related initiatives, and discuss additional considerations such as security and abuse prevention measures. We evaluate the protocol’s effectiveness by analysing multiple, independent implementations, which pass a suite of formal tests and can be demonstrated interoperating with each other. To experience the described features please open this document in your Web browser under its canonical URI: http://csarven.ca/linked-data-notifications

                        \section{Keywords}
  \label{keywords}

                            \begin{itemize}
  \item \empty Communications protocol\item \empty Decentralisation\item \empty Linked Data\item \empty Social web
    \end{itemize}

                        \section{Introduction}
  \label{introduction}

\par Notifications are sent over the Web for a variety of purposes, including social applications: ``You have been invited to a graduation party!'', ``Tim commented on your blog post!'', ``Liz tagged you in a photo''. The notification data may be displayed to a human to acknowledge, or used to trigger some other application-specific process (or both). In a decentralised architecture, notifications can be a key element for federation of information, and application integration. However in centralised systems which prevail today, this data is structured arbitrarily and typically only usable by the application that generated it in the first place. Current efforts towards {\em re-decentralising} the Web [\empty 1, \empty 2, \empty 3] are moving towards architectures in which data storage is decoupled from application logic, freeing end users to switch between applications, or to let multiple applications operate over the same data. So far, notifications are considered to be {\em ephemeral} resources which may disappear after transport, and thus are excluded from being designed for reuse.

\par We argue that notification data should not be locked into particular systems. We designed the {\em Linked Data Notifications (LDN)} protocol to support sharing and reuse of notifications {\em across} applications, regardless of how they were generated or what their contents are. We describe how the principles of identification, addressability and semantic representation can be applied to notifications on the Web. Specifying LDN as a formal protocol allows independently implemented, heterogeneous applications which generate and use notifications, to seamlessly work together. Thus, LDN supports the decentralisation of the Web as well as encourages the generation and consumption of Linked Data.

\par We build on existing W3C standards and Linked Data principles. In particular, the storage of notifications is compatible with the Linked Data Platform standard; notifications are identified by HTTP URIs; and notification contents are available as JSON-LD. A key architectural decision is the separation of concerns between {\em senders}, {\em receivers}, and {\em consumers} of notifications. Implementations of the protocol can play one or more of these roles, and interoperate successfully with implementations playing the complementary roles. This means that notifications generated by one application can be reused by a completely different application, accessed via the store where the notification data resides, through shared Linked Data vocabularies. LDN also pushes the decentralised approach further by allowing any {\em target} resource to advertise its Inbox anywhere on the Web; that is, targets do not need to be coupled to or controlled by a receiver, and can make use of a third-party {\em Inbox as a service}.

\par LDN is a W3C \empty Candidate Recommendation via the \empty Social Web Working Group [\empty 4]. The first two authors of this article are the co-editors of the specification.

\par Use cases for decentralised notifications are particularly evident in social networking (status updates, interactions, games); scholarly communication (reviews, citations); and changes of state of resources (datasets, versioning, sensor readings, experimental observations). We describe the requirements which guided the development of the protocol and discuss related work, including current alternative approaches and complementary protocols which can work alongside LDN. We summarise the protocol itself, and specific architectural considerations that were made. We built a test suite which can be used to confirm that implementations conform with the specification, and we describe 17 implementations which interoperate with each other.

\par As the following terms used throughout this article may be subject to different interpretations by different communities, we provide some definitions here.

\par By decentralisation, we mean data and applications are loosely coupled, and users are empowered to choose where their data is stored or held. We focus on Web-based decentralisation, where content is transported over HTTP, and resources are identified with URIs. An Inbox is a container or directory (attached to a Web resource) which is used to store and serve a collection of notifications. A notification is a retrievable resource which returns RDF. The contents of notifications are intended to describe a change in state of some other resource, or contain new information for the attention of a user or process, and may be subject to constraints of the Inbox it is contained in.

                        \section{Related Work}
  \label{related-work}

\par Here we review previous and ongoing efforts towards delivering notifications in a decentralised manner. Many systems which make use of notifications operate either in a completely centralised way, or are decentralised only in the sense that different instances of the {\em same} codebase need to interoperate; we restrict our review to mechanisms which do not expect the notification to be received or used only by the same software or platform which sent it.

\par The contents of a notification is either: 1) URLs, indicating relations between Web resources, or 2) a ‘fat ping’ containing a blob of information. Semantic Pingback, Webmention, and Provenance Pingback follow the first form, and are also known as linkbacks, the suite of protocols that essentially allows Web documents to automatically reciprocate hyperlinks. This has the advantage that a verification mechanism can be tightly specified (the URL of the target must appear in the content of the source), but the disadvantage that notifications are only available for use cases involving Web publishing.

\par \empty Semantic Pingback [\empty 2] and \empty Webmention [\empty 5] both update the original \empty Pingback [\empty 6] mechanism by replacing the XML-RPC transport mechanism by a {\tt x-www-form-urlencoded} request with two parameters ({\tt source} and {\tt target}). Resources which are the target for a notification advertise the respective receiving service or endpoint via a {\tt Link} relation, either in HTTP headers or HTML. Semantic Pingback additionally enables discovery of the Pingback service where target is available as RDF. While the content at source may indicate (in any convention or serialisation format) the type of relation between the source and target URLs, this information about the relation is not transmitted to the receiver’s endpoint; only the source and target URLs are sent. As such, there is also no way to distinguish between multiple potential mentions of the target at the source; this is left up to the receiver to interpret. Semantic Pingback does encourage generation of additional semantics about the relation(s) between the source and the target by processing the source as RDF if possible, and also defines specific ways for a receiving server to handle incoming pingback data in order to add the source data to an RDF knowledge base [\empty 2]. Beyond verifying that the source contains the URL of the target, Webmention does not specify any further requirements of the receiving server; nor is it expected that “mentions” are retrievable once they have been sent.

\par A \empty Provenance Pingback endpoint is also advertised via the HTTP {\tt Link} header; it accepts a list of URIs for provenance records describing uses of the resource [\empty 7]. Provenance Pingback does not specify any further behaviour by the receiving server, but the contents at the URIs listed in the notification body must be semantic data.

\par Other notification mechanisms send more information than just URLs in the notification body; due to each mechanism’s focused use case, the payload is restricted to a particular vocabulary.

\par \empty DSNotify is a centralised service which crawls datasets and observes changes to links with the specific use case of preserving link integrity between Linked Open Data resources. Third-party applications can register with the sending service to receive notifications of changes in the form of a specific XML payload [\empty 8]. With the \empty sparqlPuSH service, users may input a SPARQL query, the results of which are the specific updates they are interested in. The query is run periodically by the service, and the results are converted to RSS and Atom feeds, which is sent to a \empty PubSubHubbub hub to which the user can subscribe [\empty 9]. The \empty ResourceSync Change Notification specification also sends update notifications via a PuSH hub, this time with an XML payload based on the Sitemap format [\empty 10]. Each of these mechanisms is triggered by subscription requests. That is, a user must actively solicit messages from a particular service, rather than having a way for a service to select a notification target and autonomously discover where to send notifications to.

                        \section{Requirements and Design Considerations}
  \label{requirements-and-design-considerations}

\par In this section we discuss our considerations for a Web notification protocol that conforms to the Linked Data design principles, as well as best practices for applications. We use these considerations to establish both concrete requirements and points of implementation-specific flexibility for the protocol.

                                \subsection{R1 Modularity}
  \label{modularity}

\par To encourage modularity of applications, one should differentiate between different classes of implementation of the protocol. Two parties are involved in the creation of a notification: a {\em sender}, generating the notification data, and a {\em receiver}, storing the created resource. We also have the role of a {\em consumer}, which reads the notification data and repurposes it in some way. A software implementation can of course play two or all three of these roles; the important part is that it need not. A consuming application can read and use notification data without being concerned about ever sending or storing notifications.

                                \subsection{R2 Reusable notifications}
  \label{reusable-notifications}

\par The relationship between the {\em consumer} and {\em receiver} roles is key to notifications being reusable. A consumer must be able to autonomously find the location of notifications for or about the particular resource it is interested in. To achieve this we place a requirement on the receiver to expose notifications it has been sent in such away to permit other applications to access them; and specify how any resource can advertise its receiving endpoint for consumers to discover. To promote fair use or remixing of notification contents, applications can incorporate rights and licensing information into the data. Similarly, applications may include additional information on licensing resources that the notification refers to. The presence of this type of information is important for consumers to assess the (re)usability of data.

                                \subsection{R3 Persistence and Retrievability}
  \label{persistence-and-retrievability}

\par There is a social expectation and technical arguments for ensuring the persistence of identifiers of Web resources [\empty 11]. This is inconsistent with the traditionally ephemeral nature of notifications. Applications may benefit from referring to or reusing notifications if the notifications are known to be available in the long term, or indicate their expected lifespan [\empty 12].

\par A {\em RESTful architecture} [\empty 13] is well suited for persistent notifications, as it involves organisation of atomic resources, their discovery and description, and a lightweight API for the CRUD (create, read, update, and delete) operations [\empty 14]. This enforces the notion that notifications are considered resources in their own right, with their own dereferencable URIs.

\par We need to consider both the needs of software systems and humans when large amounts of notification data are being generated and shared between diverse applications which may be operating without knowledge of each other. To organise and manage large amount of notifications over time, mechanisms should be in place to break representations of collections of notifications into multiple paged responses that may be easier to consume by applications.

\par Relatedly, receivers may carry out resource management or garbage collection, or permit consumers or other applications to do so. For example, an application to consume messages might let an authenticated and authorised user ‘mark as read’ by adding a triple to the notification contents.

                                \subsection{R4 Adaptability}
  \label{adaptability}

\par Linked Data applications benefit from domain-driven designs; that is, functionality being small and focussed on a particular purpose, rather than generic. We believe a notification protocol should be adaptable for different domains, but that there is no need to create multiple domain-specific notification protocols; the fundamental mechanics are the same.

\par R4-A: Any resource may be the {\em target} of a notification. By target, we mean a notification may be addressed {\em to} the resource, be {\em about} the resource, or for a sender to otherwise decide that it is appropriate to draw the attention of the resource (or resource owner) to the information in the notification body. As such, any Web resource must be able to advertise an endpoint to which it can receive notifications. Resources can be RDF or non-RDF (such as an image, or CSV dataset), and may be informational (a blog post, a user profile) or non-informational (a person).

\par R4-B: We do not purport to be able to design a notifications ontology which is appropriate for every domain. Thus we consider the {\em contents} of a notification to be application specific. From a sender’s perspective, we derive two core principles: a notification can contain {\em any data}; a notification can use {\em any vocabulary}. From a consumer’s perspective, interoperability between different applications occurs through vocabulary reuse, and shared understanding of terms. This is in accordance with Linked Data principles in general. The practical upshot of this is that a calendar application which consumes event invitations using the \empty RDF Calendar vocabulary is likely to completely ignore notifications containing the \empty PROV Ontology, even if it finds them all stored in the same place. For two independent applications operating in the {\em same} domain, a shared understanding of appropriate vocabulary terms is assumed.

\par However from a receiver’s perspective, exposing itself to receive any blobs of RDF data from unknown senders may be problematic. Thus, R4-C: it should be possible for the receiver to enforce restrictions and accept only notifications that are acceptable according to its own criteria (deemed by e.g., user configuration; domain-specific receivers). This can be used as an anti-spam measure, a security protection, or for attaining application and data integrity.

\par Rejecting notifications which do not match a specific pattern in their contents, or the {\em shape} of the data, is one way to filter. For example, if the Inbox owner knows that they will only ever use a consuming application which processes friend requests, they can configure their receiver to filter out anything that does not match the pattern for a friend request, helping their consumer to be more efficient. If the notification constraints are also advertised by the receiving service as structured descriptions, generation and consumption of the notifications can be further automated. Possible specifications for doing so are W3C \empty Shapes Constraint Language (SHACL) [\empty 15] or \empty ShEx.

\par Receivers may wish to filter notifications by verifying the sender, through for example a whitelist or a Web of trust. This requires an authentication mechanism and since different authentication mechanisms are appropriate for different applications, the notification protocol should ideally be usable alongside various methods such as clientside certificates, e.g., WebID+TLS, token-based, e.g., OAuth 2.0, or digital signatures.

\par As ``anyone can say anything about anything'' a receiver may choose to resolve any external resources referred to by the notification, and cross-check the notification contents against authoritative sources. This is similar to how Semantic Pingback and Webmention require fetching and parsing of the source URL to verify existence of the target link.

                                \subsection{R5 Subscribing}
  \label{subscribing}

\par In general, applications may require that new notifications are pushed to them in real-time, or to request them at appropriate intervals. To take this into account, we expand our definition of senders, receivers and consumers with the following interaction expectations: notifications are {\em pushed} from senders to receivers; and {\em pulled} from receivers by consumers.

\par Thus, an application which offers an endpoint or callback URL to which notifications should be sent directly is a receiver, and an application which fetches notifications from an endpoint on its own schedule is a consumer. Much of the related work {\em requires} notifications to be explicitly solicited to trigger sending. Since in a decentralised model, receivers may not be aware of possible sources for notifications, our sender-receiver relationship depends on the sender’s autonomy to make such decisions by itself. This does not preclude the scenario in which a receiver may wish to solicit notifications from a particular sender, but as there are already subscription mechanisms in wide use on the Web, we do not need to specify it as part of LDN. For example, \empty WebSub (recent W3C evolution of PubSubHubbub), the WebSocket Protocol, or HTTP Web Push.

\par Given our adoption of Linked Data principles and a RESTful architecture, a further design decision was to ensure minimal compatibility with the \empty Linked Data Platform (LDP) specification [\empty 16]. LDP is a RESTful read-write API for RDF resources, which groups related resources together into constructs known as ``Containers''. Thus, existing LDP servers can be used to store notifications, as new notifications can be created by {\tt POST}ing RDF to a container.

                        \section{The LDN Protocol}
  \label{protocol}

\par The {\em Linked Data Notifications (LDN)} protocol describes how servers (receivers) can receive messages pushed to them by applications (senders), as well as how other applications (consumers) may retrieve those messages. Any resource can advertise a receiving endpoint (Inbox) for notification messages. Messages are expressed in RDF, and can contain arbitrary data. It is not dependent on a complete implementation of LDP, but comprises an easy-to-implement subset. LDN is a \empty W3C Candidate Recommendation [\empty 4].

                                Overview of Linked Data Notifications

                                \subsection{Sender to receiver interactions}
  \label{sender-to-receiver}

\par The following steps (in order without skipping) describe the interaction between sender and receiver:

\par (1) A sender is triggered, either by a human or an automatic process, to deliver a notification; (2) The sender chooses a target resource to send notifications to; (3) The sender discovers the location of the target’s {\em Inbox} through the {\tt ldp:inbox} relation in the HTTP {\tt Link} header or RDF body of the target resource; (4) The sender creates the body of the notification according to the needs of application; (5) The sender makes a {\tt POST} to the Inbox URL, containing the body in JSON-LD or in another serialisation acceptable by the server; (6) The receiver optionally applies filtering rules, and sends the appropriate HTTP response code to accept or reject the notification; (7) The receiver exposes the notification data (according to appropriate access control) for use by consumers.

                                \subsection{Consumer to receiver interactions}
  \label{consumer-to-receiver}

\par The following steps (in order without skipping) describe the interaction between consumer and receiver:

\par (1) A consumer selects a target and discovers the location of its Inbox in the same way as the sender; (2) A receiver responds to {\tt GET} requests made to the Inbox URL with a listing of the URLs of notifications that have previously been accepted, linked to the Inbox with the {\tt ldp:contains} predicate; (3) The receiver responds to {\tt GET} requests made to the individual notification URLs with JSON-LD (or optionally other serialisations); (4) Following the retrieval of notification listings or individual notifications, the consumer may perform further processing, combine with some other data, or simply present the results in a suitable human-readable way.

                                \subsection{Example Notifications}
  \label{example-notifications}

\par For more example notification payloads, see the \empty LDN specification.

                                        \begin{verbatim}{
  "@context": { "sioc": "http://rdfs.org/sioc/ns#" }
  "@id": "",
  "@type": "sioc:Comment",
  "sioc:content": "This is a great article!",
  "sioc:reply_of": { "@id": "http://example.org/article" },
  "sioc:created_at": { "@value": "2015-12-23T16:44:21Z" }
}\end{verbatim}
                                        A notification about a comment created by a user (JSON-LD).

                                        \begin{verbatim}@prefix as: <https://www.w3.org/ns/activitystreams#> .
@prefix cito: <http://purl.org/spar/cito/> .
<> a as:Announce
  as:object <https://linkedresearch.org/resources#r-903b83> ;
  as:target <http://csarven.ca/dokieli#architecture> .
<https://linkedresearch.org/resources#r-903b83>
  cito:citesAsPotentialReading
    <http://csarven.ca/linked-data-notifications#protocol> .\end{verbatim}
                                        An announcement of a specific citation relation between two entities (Turtle).

                        \section{Implementations}
  \label{implementations}

\par Here we summarise the 17 LDN implementations we are aware of to date. They are built by 10 different teams or individuals using different tool stacks (5 clientside JavaScript, 3 PHP, 3 NodeJS, 3 Python, 1 Perl, 1 Virtuoso Server Pages, 1 Java) and have submitted \empty implementation reports as part of the W3C standardisation process. We note that any \empty LDP implementation is a conforming LDN receiver; we refer here to the ones we have tested. We discuss the value of these implementations further in the \empty Evaluation section.

                                LDN Implementations

                                        Implementation
                                        Class\empty *
                                        Description

                                        \begin{description}
  \item[*] Conformance classes: S – sender, C – consumer, R – receiver.\item[a] Implementations by the authors
    \end{description}

\par Source: \empty https://github.com/w3c/ldn/tree/master/implementations

                                        \empty CarbonLDP
                                        R
                                        Data storage platform (LDP)

                                        \empty dokielia
                                        S,C
                                        Clientside editor and annotator

                                        \empty errola
                                        S
                                        Generic message sending client

                                        \empty Fedora Commons
                                        R
                                        Open source repository platform (LDP)

                                        \empty IndieAnndroid
                                        R
                                        Personal blogging platform

                                        \empty Linked Edit Rules
                                        S
                                        Statistical dataset consistency checker

                                        \empty mayktsoa
                                        R
                                        Personal data store (LDP)

                                        \empty OnScreena
                                        C
                                        Notifications display client

                                        \empty pyldn
                                        R
                                        Standalone Inbox

                                        \empty RDF-LinkedData-Notifications
                                        R
                                        Standalone Inbox

                                        \empty slopha
                                        S,R
                                        Social publishing \& quantified self

                                        \empty Solid Words
                                        S
                                        Foreign language learning app

                                        \empty solid-client
                                        S
                                        Clientside library for LDP

                                        \empty solid-inbox
                                        C
                                        Clientside social message reader

                                        \empty solid-notifications
                                        S,C
                                        Clientside library for LDN

                                        \empty solid-server
                                        R
                                        Personal data storage server (LDP)

                                        \empty Virtuoso+\empty ODS Briefcase
                                        R,C
                                        Personal data storage server (LDP)

\par We highlight social scholarly communication use cases with \empty dokieli, a clientside editor for decentralised scientific article publishing, annotations and social interactions [\empty 17]. dokieli uses LDN to send and consume notifications: When a reader comments on a fragment of text in an article, the application discovers the article’s Inbox and sends a notification about the annotation. dokieli also consumes notifications from this Inbox to fetch and display the annotation as marginalia (\empty figure 2). A reader can share a dokieli-enabled article with their contacts; dokieli discovers each contact’s Inbox and sends a notification there (\empty figure 3). When editing an article, the author can add a citation. If an Inbox is discovered in the cited article, dokieli sends a notification there to indicate what part of the article was cited by whom and where. dokieli-enabled articles also consume citation notifications to display these metrics for the author and other readers (\empty figure 4).

                                \empty Video of dokieli Web Annotation

                                \empty Video of dokieli Share

\par Notifications sent by dokieli can be reused by any consuming applications that recognise the vocabulary terms; similarly, dokieli can consume notifications sent by different applications.

\par Further social use cases are demonstrated by \empty sloph, a personal publishing and quantified self platform which acts as a node in a decentralised social network. When new content is created on the server, sloph performs discovery on URLs it finds as values of particular properties of the new content, as well as any URLs in the body of the content, and sends notifications accordingly. For instance:

                            \begin{itemize}
  \item If a {\em Like} activity is generated on the server, sloph uses the {\tt object} of the {\em Like} as the target for a notification. Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0 [\empty 18]), if the target is a dokieli article, this {\em Like} will be displayed (\empty figure 5).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e. places behind access control) any that it cannot automatically verify refer to third-party content published on another domain. If an article written with dokieli publishes a citation of a blog post which advertises a sloph Inbox, sloph will fetch the article and verify whether the relation matches the contents of the notification before exposing the notification for re-use.
    \end{itemize}

\par \empty Linked Edit Rules and \empty Solid Words are specialised senders. Linked Edit Rules checks the consistency of statistical datasets against structured constraints, and delivers the consistency report as a notification to the user. Solid Words is a clientside game for learning new words in a foreign language; it delivers the player’s score for each round to their Inbox. \empty OnScreen is a (crude) generic consumer; as such, it can display notifications sent by both of the aforementioned senders (\empty figure 6).

                                A {\em Like} notification created by sloph, displayed by dokieli.

                                A: Solid Words (a sender), B: Linked Edit Rules (a sender), C: OnScreen (a consumer) displaying notifications sent by A and B.

                        \section{Analysis and Evaluation}
  \label{analysis-and-evaluation}

\par The LDN protocol describes the discovery of a resource’s Inbox whence notifications are sent or consumed, and the sending and exposure of those notifications. Here we analyse how well features of LDN achieve the \empty requirements identified previously, and compare this to related work.

\par We have already examined \empty implementations of the specification and described how they interoperate with each other; this can be further tested by running the \empty test suite: https://linkedresearch.org/ldn/tests/. We can use this towards an evaluation of its feasibility and effectiveness at interoperability. Given the relatively early stage in the standardisation process (LDN entered Candidate Recommendation in 2016-11), the fast adoption of the LDN specification, quantity of the implementations, and their diversity is promising and further shows LDN’s feasibility. Furthermore, during the development of the specification issues have been raised or discussed by 28 different people (excluding the authors; 21 outside of the Social Web Working Group, 7 within) and the specification has undergone formal review by internationalisation, accessibility, and security specialists. We also discuss in more depth particular challenges that were raised and resolved as part of this process.

                                \subsection{Comparison summary}
  \label{comparison-summary}

\par Here we compare existing notification mechanisms from related work. The criteria includes our \empty requirements and design considerations ({\em Rx}) along with additional technical information which helps to capture some design differences ({\em Tx}).

                                        Comparison of notification mechanisms

                                                Mechanism
                                                T1
                                                T2
                                                T3
                                                R1
                                                R2
                                                R3
                                                R4-A
                                                R4-B
                                                R4-Cp
                                                R4-Cv
                                                R4-Co
                                                R5

                                                    \begin{description}
  \item[T1] Notification type\item[T2] Delivery method\item[T3] Dependencies\item[R1] Modularity (application classes: S Sender, R Receiver, C Consumer, U Subscriber/User)\item[R2] Reusability\item[R3] Persistence - required? how?\item[R4-A] Target representation\item[R4-B] Notification body\item[R4-Cp] Payload processing required?\item[R4-Cv] Verification - required? how?\item[R4-Co] Requirements for referenced resources?\item[R5] Subscription
    \end{description}

                                                    \begin{description}
  \item[–] not applicable, out of scope\item[/] not specified, in scope\item[X] explicitly disallowed\item[app] application specific decision\item[!] required ({\em MUST})\item[+] recommended ({\em SHOULD})\item[O] optional ({\em MAY})\item[PuSH] PubSubHubbub
    \end{description}

                                                    \begin{description}
  \item[h] HTML recommended\item[j] Alternate RDF formats can be negotiated\item[k] {\tt source} and {\tt target} key–value pairs is required\item[q] Provenance records with \empty PROV Ontology\item[r] RDF representation recommended\item[ra] SPARQL results transformed to RSS/Atom\item[s] \empty Sitemaps\item[t] Described in an RDF store or dataset
    \end{description}

                                                Semantic Pingback
                                                Linkback
                                                POST
                                                RDF
                                                S R
                                                /
                                                /
                                                Anyr
                                                form urlencodedk
                                                !
                                                ! parse source
                                                Anyr
                                                X

                                                Webmention
                                                Linkback
                                                POST
                                                HTML
                                                S R
                                                –
                                                –
                                                Anyh
                                                form urlencodedk
                                                !
                                                ! parse source
                                                Anyh
                                                X

                                                Provenance Pingback
                                                Linkback
                                                POST
                                                RDF
                                                S R
                                                /
                                                /
                                                /
                                                URI list
                                                /
                                                /
                                                RDFq
                                                X

                                                DSNotify
                                                Fat ping
                                                POST, PUT
                                                XML, PuSH
                                                S U
                                                /
                                                –
                                                –
                                                XML
                                                /
                                                –
                                                RDFt
                                                !

                                                sparqlPuSH
                                                Fat ping
                                                POST
                                                XML, SPARQL, PuSH
                                                S U
                                                –
                                                –
                                                –
                                                XMLra
                                                /
                                                –
                                                RDFt
                                                !

                                                ResourceSync
                                                Fat ping
                                                POST
                                                XML, PuSH
                                                S U
                                                /
                                                –
                                                –
                                                XMLs
                                                /
                                                –
                                                ?
                                                !

                                                Linked Data Notifications
                                                Fat ping
                                                POST
                                                JSON-LD
                                                S R C
                                                !
                                                ! URI
                                                Any
                                                JSON-LDj
                                                + app
                                                + app
                                                –
                                                O app

\par Given that each application requires to follow the steps listed in ``\empty Sender to receiver interaction'' and ``\empty Consumer to receiver interactions'' the metrics are dependent on the performance of client and server to do HTTP requests and responses, and their respective payloads.

                                \subsection{Compatibility with existing systems}
  \label{compatibility-with-existing-systems}

\par Per \empty R1 and \empty R4 we have tried to optimise LDN for use as a module of a larger system. The success of this is demonstrated by implementations which use LDN alongside existing protocols according to their specific needs.

\par The Solid suite of tools, Virtuoso+ODS-Briefcase, and dokieli use \empty Web Access Control along with an authentication mechanism to apply fine grained access controls to restrict who can send notifications, or who can retrieve notifications from the Inbox. sloph demonstrates an Inbox as a \empty Webhooks callback URL, for requesting notifications from APIs which post JSON-based payloads. \empty ActivityPub is a W3C CR for decentralised social media [\empty 19]. It uses LDN for delivery of notifications with the \empty ActivityStreams 2.0 (AS2) vocabulary, and specifies additional specialised receiver behaviour; also used by sloph. dokieli uses the \empty Web Annotation Protocol, an LDP-based mechanism for creating new content, which acts as a trigger for notifications to be sent to the Inbox of the annotation target. The \empty Fedora API Specification is in the process of being formalised (as an extension of LDP) by the Fedora community. The repository event stream draws upon the LDN specification, allowing LDN consumers and senders to react asynchronously to repository events.

\par Any existing LDP implementation can serve as an LDN receiver. Simply advertising any {\tt ldp:Container} as the Inbox for a resource is sufficient. We confirmed this with four LDP servers which were developed independently with different code bases, prior to the LDN specification (CarbonLDP, Fedora Commons, Solid Server, Virtuoso).

\par LDN has been integrated into existing domain specific systems: dokieli, Fedora Commons, IndieAnndroid, Linked Edit Rules, sloph, solid-client, Solid Words. Standalone implementations of LDN are also straightforward as a result of this modularity, ie: errol, mayktso, onscreen, pyLDN, RDF-LinkedData-Notifications, solid-inbox, solid-notifications.

                                \subsection{Optimising implementation}
  \label{optimising-implementation}

\par We have considered tradeoffs between the HTTP operations receivers and publishers are {\em required} to respond to, and ways in which developers may wish to optimise senders or consumers by reducing outbound requests.

\par

                                        {\tt HEAD} requests are low cost, and {\tt GET} requests may be high cost if the body of the resource is large.

                                        Given that an Inbox may be discovered from the HTTP headers of a resource, senders and consumers can optimise by attempting a {\tt HEAD} request for discovery, and only continuing with a {\tt GET} request if the {\tt HEAD} is not successful. On the other hand, senders and consumers may be attempting discovery upon RDF resources which they already intend to parse into their own storage. In this case, there is no need for a {\tt HEAD} request, as a {\tt GET} will yield both HTTP {\tt Link} headers and an RDF body, either of which could include the Inbox triple. This means that resources advertising an Inbox must respond to {\tt GET} requests (even if only with HTTP headers) and may respond to {\tt HEAD} requests.

                                \subsection{Data Formats and Content Negotiation}
  \label{data-formats}

\par Handling data irrespective of the particular RDF serialisation permits some flexibility, but can be costly to support. We take into account: (a) application interoperability, (b) maintenance of RDF parsers and serialisation libraries, (c) complexity of their inclusion in applications, (d) run-time efficiency.

\par

                                        To address these issues, LDN requires all applications to create and understand the JSON-LD syntax, both for the contents of Inbox as well as for individual notifications. Choosing a single serialisation to {\em require} is necessary for consistent interoperability, as well as keeping processing requirements or external code dependencies minimal.

                                        JSON-LD is advantageous in being familiar for developers who are \empty used to JSON-based APIs but not RDF [\empty 20], and it is compatible with existing JSON libraries or in some cases native programming language data structures.

\par Optionally, applications may attempt to exchange different RDF serialisations by performing content negotiation (receivers can expose {\tt Accept-Post} headers for senders, and consumers can send {\tt Accept} headers to receivers).

                                \subsection{Precision}
  \label{precision}

\par In placing no constraints on the contained information, LDN enables a sender to be precise and lossless with the data it is transmitting. Approaches which send only URLs rely on the receiver interpreting a third-party resource, which may or may not contain structured markup or be under the control of the sender. Approaches which offer additional guidance to aid the receiver in interpreting the source document(s) nonetheless still restricts the sender. LDN therefore offers flexibility to senders, increasing the potential uses for the notification mechanism. LDN compensates for increased complexity on the receiver’s end by recommending filtering mechanisms, and moving some of the burden of understanding notifications to the consumer role. As such LDN can cover a broader variety of use cases.

                                \subsection{Accommodating different targets}
  \label{accommodating-different-targets}

\par Per {\em R4 Adaptability}, we want LDN to be available for all resources in any publishing context. We consider lowering the bar for publishers of target resources to be a worthwhile trade-off against slightly increased complexity for senders and consumers. This is why we require that senders and consumers must be equipped to discover Inboxes through both HTTP headers and RDF content.

\par Since binary formats such as images and video cannot contain an RDF relation, the HTTP header is essential for including them. It also allows the inclusion of resources for which it is undesirable or impractical to add individual Inbox relations, such as to elements in a dataset; or circumstances where the developer responsible for the Inbox relation is unable to modify the content. Conversely, non-informational resources (represented with fragment URIs or 303 redirects) are unable to express HTTP headers. Their relation to an Inbox must be expressed in an RDF source. However, if a sender or consumer has a domain-specific requirement to {\em only} ever target non-informational resources, they are exempt from the requirement of discovery via HTTP headers.

                        \section{Conclusions}
  \label{conclusions}

\par In this article we describe LDN, a protocol for decentralised semantic notifications, currently undergoing standardisation at the W3C. Key elements are:

                            \begin{itemize}
  \item Notifications as retrievable, reusable entities with their own URIs.\item Distinct conformance classes for senders, receivers, and consumers.\item Deliberately not defining the vocabulary of notification contents to allow for use in a range of different application domains.\item Flexibility of authentication and verification, for the same reason.
    \end{itemize}

\par We outlined design requirements, describe how LDN meets these, and compare this with related work. We consider LDN to have greater modularity and adaptability to different scenarios, as well as good conformance with Linked Data principles. This specification has potential to have high impact in increasing interoperability between decentralised Linked Data applications in related domains, as well as generating new discoverable content for the LOD Cloud. This is evidenced by 17 diverse implementations which can be shown to interoperate with each other, including generic libraries and datastores, and domain-specific applications. Being on the W3C standards track increases the likelihood of further adoption.

                            \begin{enumerate}
  \item Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016, \empty http://www2016.net/proceedings/companion/p223.pdf\item Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012, \empty \item Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016, \empty http://events.linkeddata.org/ldow2016/papers/LDOW2016\_paper\_02.pdf\item Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/ldn/\item Parecki, A.: Webmention, W3C Proposed Recommendation, 2016, \empty https://www.w3.org/TR/webmention/\item Langridge, S., Hickson, I.: Pingback 1.0, 2002, \empty http://www.hixie.ch/specs/pingback/pingback\item Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \empty http://www.w3.org/TR/prov-aq/\item Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010, \empty http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf\item Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol. 699, 2010, \empty http://ceur-ws.org/Vol-699/Paper6.pdf\item Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016, \empty http://www.openarchives.org/rs/notification/1.0/notification\item Berners-Lee, T.: Cool URIs don't change, W3C, 1998, \empty https://www.w3.org/Provider/Style/URI.html\item Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On Persistent URIs, 2012, \empty http://philarcher.org/diary/2013/uripersistence/\item Fielding, R. T.: Architectural Styles and the Design of Network-based Software Architectures. Doctoral dissertation, University of California, Irvine, 2000, \empty http://www.ics.uci.edu/\~fielding/pubs/dissertation/rest\_arch\_style.htm\item Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011, \empty http://ws-rest.org/2011/proc/a5-page.pdf\item Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016, \empty \item Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C Recommendation, 2015, \empty https://www.w3.org/TR/ldp/\item Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016, \empty http://csarven.ca/dokieli\item Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitystreams-core/\item Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitypub/\item Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014, \empty http://manu.sporny.org/2014/json-ld-origins-2/
    \end{enumerate}

    \end{document}

### 2017-03-07 index.tex
Version that compiles without errors. Not pretty output

@@
+\documentclass[a4paper]{llncs}
+\usepackage{ifxetex}
+\ifxetex
+\usepackage{fontspec}
+\defaultfontfeatures{Ligatures=TeX}
+\else
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\fi
+
+\usepackage{listings}
+\lstset{numbers=left, captionpos=b, frame=single, basicstyle=\ttfamily, stringstyle=\small, breaklines=true, extendedchars=true, showstringspaces=true, numberbychapter=false}
+
+\usepackage{caption}
+\captionsetup{labelfont=bf}
+
+\usepackage{verbatim}
+
+\usepackage[numbers]{natbib}
+\usepackage{dcolumn}
+\usepackage[export]{adjustbox}
+\usepackage{graphicx}
+\usepackage[multidot]{grffile}
+\usepackage{url}
+\usepackage{xspace}
+\usepackage{color}
+\usepackage{float}
+\usepackage{multirow}
+\usepackage{booktabs}
+\usepackage{tabularx}
+\usepackage{array}
+\usepackage{todonotes}
+\usepackage{hyperref}
+\usepackage{enumitem}
+\setlist[itemize,enumerate]{leftmargin=*}
+
+\def\sectionautorefname{Section}
+%\renewcommand{\bibname}{References}
+
   \begin{document}
@@
                     \begin{description}
-  \item[Authors] \empty Sarven Capadisli\empty 1\empty ✊\empty Amy Guy\empty 2\empty 🐦\empty Christoph Lange\empty 1\empty ∫\empty Sören Auer\empty 1\empty ⚛\empty Tim Berners-Lee\empty 3\empty 🕸
+  \item[Authors] \empty Sarven Capadisli\empty 1\empty a\empty Amy Guy\empty 2\empty b\empty Christoph Lange\empty 1\empty c\empty Sören Auer\empty 1\empty d\empty Tim Berners-Lee\empty 3\empty e
     \end{description}
@@
                     \begin{itemize}
-  \item ✊\empty info@csarven.ca\item 🐦\empty amy@rhiaro.co.uk\item ∫\empty langec@cs.uni-bonn.de\item ⚛\empty auer@cs.uni-bonn.de\item 🕸\empty timbl@w3.org
+  \item a\empty info@csarven.ca\item b\empty amy@rhiaro.co.uk\item c\empty langec@cs.uni-bonn.de\item d\empty auer@cs.uni-bonn.de\item e\empty timbl@w3.org
     \end{itemize}

### 2017-03-07 index.tex
Sync

@@
                     \begin{itemize}
-  \item a\empty info@csarven.ca\item b\empty amy@rhiaro.co.uk\item c\empty langec@cs.uni-bonn.de\item d\empty auer@cs.uni-bonn.de\item e\empty timbl@w3.org
+  \item ✊\empty info@csarven.ca\item 🐦\empty amy@rhiaro.co.uk\item ∫\empty langec@cs.uni-bonn.de\item ⚛\empty auer@cs.uni-bonn.de\item 🕸\empty timbl@w3.org
     \end{itemize}
-
-                \begin{description}
-  \item[Identifier] \empty http://csarven.ca/linked-data-notifications
-    \end{description}
 

### 2017-03-07 index.tex
TeX link

@@
 We evaluate the protocol’s effectiveness by analysing multiple, independent implementations, which pass a suite of formal tests and can be demonstrated interoperating with each other.
-To experience the described features please open this document in your Web browser under its canonical URI: http://csarven.ca/linked-data-notifications
+To experience the described features please open this document in your Web browser under its canonical URI: {\tt \href{http://csarven.ca/this-paper-is-a-demo}{http://csarven.ca/this-paper-is-a-demo}}.
 

### 2017-03-07 index.tex
Link to canonical URI

--- 2017-03-07 index.tex
@@
 We evaluate the protocol’s effectiveness by analysing multiple, independent implementations, which pass a suite of formal tests and can be demonstrated interoperating with each other.
-To experience the described features please open this document in your Web browser under its canonical URI: http://csarven.ca/linked-data-notifications
+To experience the described features please open this document in your Web browser under its canonical URI: {\tt \href{http://csarven.ca/linked-data-notifications}{http://csarven.ca/linked-data-notifications}}.
 

### 2017-03-07 index.pandoc.tex
Added pandoc conversion

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Linked Data Notifications: a resource-centric communication protocol},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{Linked Data Notifications: a resource-centric communication protocol}
\date{}

\begin{document}
\maketitle

\section{Linked Data Notifications: a resource-centric communication
protocol}\label{linked-data-notifications-a-resource-centric-communication-protocol}

\hypertarget{authors}{}
\begin{description}
\tightlist
\item[Authors]
{{\href{http://csarven.ca/}{{{Sarven}
{Capadisli}}}}}\textsuperscript{\protect\hyperlink{author-org-1}{1}\protect\hyperlink{author-email-1}{✊}}

{{\href{https://rhiaro.co.uk/}{{{Amy}
{Guy}}}}}\textsuperscript{\protect\hyperlink{author-org-2}{2}\protect\hyperlink{author-email-2}{🐦}}

{\href{https://langec.wordpress.com/about/}{Christoph
Lange}}\textsuperscript{\protect\hyperlink{author-org-1}{1}\protect\hyperlink{author-email-3}{∫}}

{\href{http://eis.iai.uni-bonn.de/SoerenAuer.html}{Sören
Auer}}\textsuperscript{\protect\hyperlink{author-org-1}{1}\protect\hyperlink{author-email-4}{⚛}}

{\href{https://www.w3.org/People/Berners-Lee/}{Tim
Berners-Lee}}\textsuperscript{\protect\hyperlink{author-org-4}{3}\protect\hyperlink{author-email-5}{🕸}}
\end{description}

\begin{itemize}
\item
  \hypertarget{author-org-1}{}

  \textsuperscript{1}Enterprise Information Systems Department,
  \href{http://uni-bonn.de/}{University of Bonn}, Bonn, DE
\item
  \hypertarget{author-org-2}{}

  \textsuperscript{2}School of Informatics,
  \href{http://inf.ed.ac.uk/}{University of Edinburgh}, Edinburgh, UK
\item
  \hypertarget{author-org-3}{}

  \textsuperscript{3}Decentralized Information Group, CSAIL,
  \href{https://mit.edu/}{MIT}, Cambridge, US
\end{itemize}

\begin{itemize}
\item
  \hypertarget{author-email-1}{}

  \textsuperscript{✊}\href{mailto:info@csarven.ca}{\nolinkurl{info@csarven.ca}}
\item
  \hypertarget{author-email-2}{}

  \textsuperscript{🐦}\href{mailto:amy@rhiaro.co.uk}{\nolinkurl{amy@rhiaro.co.uk}}
\item
  \hypertarget{author-email-3}{}

  \textsuperscript{∫}\href{mailto:langec@cs.uni-bonn.de}{\nolinkurl{langec@cs.uni-bonn.de}}
\item
  \hypertarget{author-email-4}{}

  \textsuperscript{⚛}\href{mailto:auer@cs.uni-bonn.de}{\nolinkurl{auer@cs.uni-bonn.de}}
\item
  \hypertarget{author-email-5}{}

  \textsuperscript{🕸}\href{mailto:timbl@w3.org}{\nolinkurl{timbl@w3.org}}
\end{itemize}

\hypertarget{content}{}
\hypertarget{abstract}{}
\subsection{Abstract}\label{abstract}

In this article we describe the Linked Data Notifications (LDN)
protocol, which is a \href{https://www.w3.org/TR/ldn/}{W3C Candidate
Recommendation}. Notifications are sent over the Web for a variety of
purposes, for example, by social applications. The information contained
within a notification is structured arbitrarily, and typically only
usable by the application which generated it in the first place. In the
spirit of Linked Data, we propose that notifications should be reusable
by multiple authorised applications. Through separating the concepts of
\emph{senders}, \emph{receivers} and \emph{consumers} of notifications,
and leveraging Linked Data principles of shared vocabularies and URIs,
LDN provides a building block for decentralised Web applications. This
permits end users more freedom to switch between the online tools they
use, as well as generating greater value when notifications from
different sources can be used in combination. We situate LDN alongside
related initiatives, and discuss additional considerations such as
security and abuse prevention measures. We evaluate the protocol's
effectiveness by analysing multiple, independent implementations, which
pass a suite of formal tests and can be demonstrated interoperating with
each other. To experience the described features please open this
document in your Web browser under its canonical URI:
http://csarven.ca/linked-data-notifications

\hypertarget{keywords}{}
\subsection{Keywords}\label{keywords}

\begin{itemize}
\tightlist
\item
  \href{https://en.wikipedia.org/wiki/Communications_protocol}{Communications
  protocol}
\item
  \href{https://en.wikipedia.org/wiki/Decentralization}{Decentralisation}
\item
  \href{https://en.wikipedia.org/wiki/Linked_data}{Linked Data}
\item
  \href{https://en.wikipedia.org/wiki/Social_web}{Social web}
\end{itemize}

\hypertarget{introduction}{}
\subsection{Introduction}\label{introduction}

Notifications are sent over the Web for a variety of purposes, including
social applications: ``You have been invited to a graduation party!'',
``Tim commented on your blog post!'', ``Liz tagged you in a photo''. The
notification data may be displayed to a human to acknowledge, or used to
trigger some other application-specific process (or both).
\protect\hypertarget{issue}{}{In a decentralised architecture,
notifications can be a key element for federation of information, and
application integration. However in {centralised systems which prevail
today}, this data is structured arbitrarily and typically only usable by
the application that generated it in the first place. Current efforts
towards \emph{re-decentralising} the Web
{[}\protect\hyperlink{ref-1}{1}, \protect\hyperlink{ref-2}{2},
\protect\hyperlink{ref-3}{3}{]} are moving towards architectures in
which data storage is decoupled from application logic, freeing end
users to switch between applications, or to let multiple applications
operate over the same data. So far, notifications are considered to be
\emph{ephemeral} resources which may disappear after transport, and thus
are excluded from being designed for reuse.}

We argue that notification data should not be locked into particular
systems. We designed the \emph{Linked Data Notifications (LDN)} protocol
to support sharing and reuse of notifications \emph{across}
applications, regardless of how they were generated or what their
contents are. We describe how the principles of identification,
addressability and semantic representation can be applied to
notifications on the Web. Specifying LDN as a formal protocol allows
independently implemented, heterogeneous applications which generate and
use notifications, to seamlessly work together. Thus, LDN supports the
decentralisation of the Web as well as encourages the generation and
consumption of Linked Data.

We build on existing W3C standards and Linked Data principles. In
particular, the storage of notifications is compatible with the Linked
Data Platform standard; notifications are identified by HTTP URIs; and
notification contents are available as JSON-LD. A key architectural
decision is the separation of concerns between \emph{senders},
\emph{receivers}, and \emph{consumers} of notifications. Implementations
of the protocol can play one or more of these roles, and interoperate
successfully with implementations playing the complementary roles. This
means that notifications generated by one application can be reused by a
completely different application, accessed via the store where the
notification data resides, through shared Linked Data vocabularies. LDN
also pushes the decentralised approach further by allowing any
\emph{target} resource to advertise its Inbox anywhere on the Web; that
is, targets do not need to be coupled to or controlled by a receiver,
and can make use of a third-party \emph{Inbox as a service}.{}

LDN is a W3C \href{https://www.w3.org/TR/ldn/}{Candidate Recommendation}
via the \href{https://www.w3.org/wiki/Socialwg}{Social Web Working
Group} {[}\protect\hyperlink{ref-4}{4}{]}. The first two authors of this
article are the co-editors of the specification.

Use cases for decentralised notifications are particularly evident in
social networking (status updates, interactions, games); scholarly
communication (reviews, citations); and changes of state of resources
(datasets, versioning, sensor readings, experimental observations). We
describe the requirements which guided the development of the protocol
and discuss related work, including current alternative approaches and
complementary protocols which can work alongside LDN. We summarise the
protocol itself, and specific architectural considerations that were
made. We built a test suite which can be used to confirm that
implementations conform with the specification, and we describe 17
implementations which interoperate with each other.

\hypertarget{concept-scheme}{}
{As the following terms used throughout this article may be subject to
different interpretations by different communities, we provide some
definitions here.}

By \textbf{decentralisation}, we mean {data and applications are loosely
coupled, and users are empowered to choose where their data is stored or
held. We focus on Web-based decentralisation, where content is
transported over HTTP, and resources are identified with URIs.} An
\textbf{Inbox} is {a container or directory (attached to a Web resource)
which is used to store and serve a collection of notifications.} A
\textbf{notification} is {a retrievable resource which returns RDF. The
contents of notifications are intended to describe a change in state of
some other resource, or contain new information for the attention of a
user or process, and may be subject to constraints of the Inbox it is
contained in.}

\hypertarget{related-work}{}
\subsection{Related Work}\label{related-work}

Here we review previous and ongoing efforts towards delivering
notifications in a decentralised manner. Many systems which make use of
notifications operate either in a completely centralised way, or are
decentralised only in the sense that different instances of the
\emph{same} codebase need to interoperate; we restrict our review to
mechanisms which do not expect the notification to be received or used
only by the same software or platform which sent it.

The contents of a notification is either: 1) URLs, indicating relations
between Web resources, or 2) a `fat ping' containing a blob of
information. Semantic Pingback, Webmention, and Provenance Pingback
follow the first form, and are also known as linkbacks, the suite of
protocols that essentially allows Web documents to automatically
reciprocate hyperlinks. This has the advantage that a verification
mechanism can be tightly specified (the URL of the target must appear in
the content of the source), but the disadvantage that notifications are
only available for use cases involving Web publishing.

\href{https://aksw.github.io/SemanticPingback/}{Semantic Pingback}
{[}\protect\hyperlink{ref-2}{2}{]} and
\href{https://www.w3.org/TR/webmention}{Webmention}
{[}\protect\hyperlink{ref-5}{5}{]} both update the original
\href{http://www.hixie.ch/specs/pingback/pingback}{Pingback}
{[}\protect\hyperlink{ref-6}{6}{]} mechanism by replacing the XML-RPC
transport mechanism by a \texttt{x-www-form-urlencoded} request with two
parameters (\texttt{source} and \texttt{target}). Resources which are
the target for a notification advertise the respective receiving service
or endpoint via a \texttt{Link} relation, either in HTTP headers or
HTML. Semantic Pingback additionally enables discovery of the Pingback
service where target is available as RDF. While the content at source
may indicate (in any convention or serialisation format) the type of
relation between the source and target URLs, this information about the
relation is not transmitted to the receiver's endpoint; only the source
and target URLs are sent. As such, there is also no way to distinguish
between multiple potential mentions of the target at the source; this is
left up to the receiver to interpret. Semantic Pingback does encourage
generation of additional semantics about the relation(s) between the
source and the target by processing the source as RDF if possible, and
also defines specific ways for a receiving server to handle incoming
pingback data in order to add the source data to an RDF knowledge base
{[}\protect\hyperlink{ref-2}{2}{]}. Beyond verifying that the source
contains the URL of the target, Webmention does not specify any further
requirements of the receiving server; nor is it expected that
``mentions'' are retrievable once they have been sent.

A \href{http://www.w3.org/TR/prov-aq/\#provenance-pingback}{Provenance
Pingback} endpoint is also advertised via the HTTP \texttt{Link} header;
it accepts a list of URIs for provenance records describing uses of the
resource {[}\protect\hyperlink{ref-7}{7}{]}. Provenance Pingback does
not specify any further behaviour by the receiving server, but the
contents at the URIs listed in the notification body must be semantic
data.

Other notification mechanisms send more information than just URLs in
the notification body; due to each mechanism's focused use case, the
payload is restricted to a particular vocabulary.

\href{http://www.cibiv.at/~niko/dsnotify/}{DSNotify} is a centralised
service which crawls datasets and observes changes to links with the
specific use case of preserving link integrity between Linked Open Data
resources. Third-party applications can register with the sending
service to receive notifications of changes in the form of a specific
XML payload {[}\protect\hyperlink{ref-8}{8}{]}. With the
\href{https://www.w3.org/2001/sw/wiki/SparqlPuSH}{sparqlPuSH} service,
users may input a SPARQL query, the results of which are the specific
updates they are interested in. The query is run periodically by the
service, and the results are converted to RSS and Atom feeds, which is
sent to a
\href{http://pubsubhubbub.github.io/PubSubHubbub/pubsubhubbub-core-0.4.html}{PubSubHubbub}
hub to which the user can subscribe {[}\protect\hyperlink{ref-9}{9}{]}.
The
\href{http://www.openarchives.org/rs/notification/1.0/notification}{ResourceSync
Change Notification} specification also sends update notifications via a
PuSH hub, this time with an XML payload based on the Sitemap format
{[}\protect\hyperlink{ref-10}{10}{]}. Each of these mechanisms is
triggered by subscription requests. That is, a user must actively
solicit messages from a particular service, rather than having a way for
a service to select a notification target and autonomously discover
where to send notifications to.

\hypertarget{requirements-and-design-considerations}{}
\hypertarget{requirements-and-design-considerations}{\subsection{Requirements
and Design
Considerations}\label{requirements-and-design-considerations}}

In this section we discuss our considerations for a {Web notification
protocol that conforms to the Linked Data design principles}, as well as
{best practices for applications}. We use these considerations to
establish both concrete requirements and points of
implementation-specific flexibility for the protocol.

\hypertarget{modularity}{}
\subsubsection{R1 Modularity}\label{r1-modularity}

To encourage modularity of applications, one should differentiate
between different classes of implementation of the protocol. Two parties
are involved in the creation of a notification: a \emph{sender},
generating the notification data, and a \emph{receiver}, storing the
created resource. We also have the role of a \emph{consumer}, which
reads the notification data and repurposes it in some way. A software
implementation can of course play two or all three of these roles; the
important part is that it need not. A consuming application can read and
use notification data without being concerned about ever sending or
storing notifications.

\hypertarget{reusable-notifications}{}
\subsubsection{R2 Reusable
notifications}\label{r2-reusable-notifications}

The relationship between the \emph{consumer} and \emph{receiver} roles
is key to notifications being reusable. A consumer must be able to
autonomously find the location of notifications for or about the
particular resource it is interested in. To achieve this we place a
requirement on the receiver to expose notifications it has been sent in
such away to permit other applications to access them; and specify how
any resource can advertise its receiving endpoint for consumers to
discover. To promote fair use or remixing of notification contents,
applications can incorporate rights and licensing information into the
data. Similarly, applications may include additional information on
licensing resources that the notification refers to. The presence of
this type of information is important for consumers to assess the
(re)usability of data.

\hypertarget{persistence-and-retrievability}{}
\subsubsection{R3 Persistence and
Retrievability}\label{r3-persistence-and-retrievability}

There is a social expectation and technical arguments for ensuring the
persistence of identifiers of Web resources
{[}\protect\hyperlink{ref-11}{11}{]}. This is inconsistent with the
traditionally ephemeral nature of notifications. Applications may
benefit from referring to or reusing notifications if the notifications
are known to be available in the long term, or indicate their expected
lifespan {[}\protect\hyperlink{ref-12}{12}{]}.

A \emph{RESTful architecture} {[}\protect\hyperlink{ref-13}{13}{]} is
well suited for persistent notifications, as it involves organisation of
atomic resources, their discovery and description, and a lightweight API
for the CRUD (create, read, update, and delete) operations
{[}\protect\hyperlink{ref-14}{14}{]}. This enforces the notion that
notifications are considered resources in their own right, with their
own dereferencable URIs.

We need to consider both the needs of software systems and humans when
large amounts of notification data are being generated and shared
between diverse applications which may be operating without knowledge of
each other. To organise and manage large amount of notifications over
time, mechanisms should be in place to break representations of
collections of notifications into multiple paged responses that may be
easier to consume by applications.

Relatedly, receivers may carry out resource management or garbage
collection, or permit consumers or other applications to do so. For
example, an application to consume messages might let an authenticated
and authorised user `mark as read' by adding a triple to the
notification contents.

\hypertarget{adaptability}{}
\subsubsection{R4 Adaptability}\label{r4-adaptability}

Linked Data applications benefit from domain-driven designs; that is,
functionality being small and focussed on a particular purpose, rather
than generic. We believe a notification protocol should be adaptable for
different domains, but that there is no need to create multiple
domain-specific notification protocols; the fundamental mechanics are
the same.

\textbf{R4-A}: Any resource may be the \emph{target} of a notification.
By target, we mean a notification may be addressed \emph{to} the
resource, be \emph{about} the resource, or for a sender to otherwise
decide that it is appropriate to draw the attention of the resource (or
resource owner) to the information in the notification body. As such,
any Web resource must be able to advertise an endpoint to which it can
receive notifications. Resources can be RDF or non-RDF (such as an
image, or CSV dataset), and may be informational (a blog post, a user
profile) or non-informational (a person).

\textbf{R4-B}: We do not purport to be able to design a notifications
ontology which is appropriate for every domain. Thus we consider the
\emph{contents} of a notification to be application specific. From a
sender's perspective, we derive two core principles: a notification can
contain \emph{any data}; a notification can use \emph{any vocabulary}.
From a consumer's perspective, interoperability between different
applications occurs through vocabulary reuse, and shared understanding
of terms. This is in accordance with Linked Data principles in general.
The practical upshot of this is that a calendar application which
consumes event invitations using the
\href{https://www.w3.org/TR/rdfcal/}{RDF Calendar} vocabulary is likely
to completely ignore notifications containing the
\href{www.w3.org/TR/prov-o/}{PROV Ontology}, even if it finds them all
stored in the same place. For two independent applications operating in
the \emph{same} domain, a shared understanding of appropriate vocabulary
terms is assumed.

\hypertarget{notification-verification}{}
However from a receiver's perspective, exposing itself to receive any
blobs of RDF data from unknown senders may be problematic. Thus,
\textbf{R4-C}: it should be possible for the receiver to enforce
restrictions and accept only notifications that are acceptable according
to its own criteria (deemed by e.g., user configuration; domain-specific
receivers). This can be used as an anti-spam measure, a security
protection, or for attaining application and data integrity.

Rejecting notifications which do not match a specific pattern in their
contents, or the \emph{shape} of the data, is one way to filter. For
example, if the Inbox owner knows that they will only ever use a
consuming application which processes friend requests, they can
configure their receiver to filter out anything that does not match the
pattern for a friend request, helping their consumer to be more
efficient. If the notification constraints are also advertised by the
receiving service as structured descriptions, generation and consumption
of the notifications can be further automated. Possible specifications
for doing so are W3C \href{https://www.w3.org/TR/shacl/}{Shapes
Constraint Language (SHACL)} {[}\protect\hyperlink{ref-15}{15}{]} or
\href{https://shexspec.github.io/spec/}{ShEx}.

Receivers may wish to filter notifications by verifying the sender,
through for example a whitelist or a Web of trust. This requires an
authentication mechanism and since different authentication mechanisms
are appropriate for different applications, the notification protocol
should ideally be usable alongside various methods such as clientside
certificates, e.g., WebID+TLS, token-based, e.g., OAuth 2.0, or digital
signatures.

As ``anyone can say anything about anything'' a receiver may choose to
resolve any external resources referred to by the notification, and
cross-check the notification contents against authoritative sources.
This is similar to how Semantic Pingback and Webmention require fetching
and parsing of the source URL to verify existence of the target link.

\hypertarget{subscribing}{}
\subsubsection{R5 Subscribing}\label{r5-subscribing}

In general, applications may require that new notifications are pushed
to them in real-time, or to request them at appropriate intervals. To
take this into account, we expand our definition of senders, receivers
and consumers with the following interaction expectations: notifications
are \emph{pushed} from senders to receivers; and \emph{pulled} from
receivers by consumers.

Thus, an application which offers an endpoint or callback URL to which
notifications should be sent directly is a receiver, and an application
which fetches notifications from an endpoint on its own schedule is a
consumer. Much of the related work \emph{requires} notifications to be
explicitly solicited to trigger sending. Since in a decentralised model,
receivers may not be aware of possible sources for notifications, our
sender-receiver relationship depends on the sender's autonomy to make
such decisions by itself. This does not preclude the scenario in which a
receiver may wish to solicit notifications from a particular sender, but
as there are already subscription mechanisms in wide use on the Web, we
do not need to specify it as part of LDN. For example,
\href{https://www.w3.org/TR/websub/}{WebSub} (recent W3C evolution of
PubSubHubbub), the WebSocket Protocol, or HTTP Web Push.

Given our adoption of Linked Data principles and a RESTful architecture,
a further design decision was to ensure minimal compatibility with the
\href{https://www.w3.org/TR/ldp/}{Linked Data Platform} (LDP)
specification {[}\protect\hyperlink{ref-16}{16}{]}. LDP is a RESTful
read-write API for RDF resources, which groups related resources
together into constructs known as ``Containers''. Thus, existing LDP
servers can be used to store notifications, as new notifications can be
created by \texttt{POST}ing RDF to a container.

\hypertarget{protocol}{}
\subsection{The LDN Protocol}\label{the-ldn-protocol}

The Linked Data Notifications (LDN) protocol describes how servers
(receivers) can receive messages pushed to them by applications
(senders), as well as how other applications (consumers) may retrieve
those messages. Any resource can advertise a receiving endpoint (Inbox)
for notification messages. Messages are expressed in RDF, and can
contain arbitrary data. It is not dependent on a complete implementation
of LDP, but comprises an easy-to-implement subset. LDN is a
\href{https://www.w3.org/TR/ldn}{W3C Candidate Recommendation}
{[}\protect\hyperlink{ref-4}{4}{]}.

Overview of Linked Data Notifications

\hypertarget{sender-to-receiver}{}
\subsubsection{Sender to receiver
interactions}\label{sender-to-receiver-interactions}

The following steps (in order without skipping) describe the interaction
between sender and receiver:

(1) A sender is triggered, either by a human or an automatic process, to
deliver a notification; (2) The sender chooses a target resource to send
notifications to; (3) The sender discovers the location of the target's
\emph{Inbox} through the \texttt{ldp:inbox} relation in the HTTP
\texttt{Link} header or RDF body of the target resource; (4) The sender
creates the body of the notification according to the needs of
application; (5) The sender makes a \texttt{POST} to the Inbox URL,
containing the body in JSON-LD or in another serialisation acceptable by
the server; (6) The receiver optionally applies filtering rules, and
sends the appropriate HTTP response code to accept or reject the
notification; (7) The receiver exposes the notification data (according
to appropriate access control) for use by consumers.

\hypertarget{consumer-to-receiver}{}
\subsubsection{Consumer to receiver
interactions}\label{consumer-to-receiver-interactions}

The following steps (in order without skipping) describe the interaction
between consumer and receiver:

(1) A consumer selects a target and discovers the location of its Inbox
in the same way as the sender; (2) A receiver responds to \texttt{GET}
requests made to the Inbox URL with a listing of the URLs of
notifications that have previously been accepted, linked to the Inbox
with the \texttt{ldp:contains} predicate; {(3) The receiver responds to
\texttt{GET} requests made to the individual notification URLs with
JSON-LD (or optionally other serialisations);} {(4) Following the
retrieval of notification listings or individual notifications, the
consumer may perform further processing, combine with some other data,
or simply present the results in a suitable human-readable way.}

\hypertarget{example-notifications}{}
\subsubsection{Example Notifications}\label{example-notifications}

For more example notification payloads, see the
\href{https://www.w3.org/TR/ldn/}{LDN specification}.

\begin{verbatim}
{
  "@context": { "sioc": "http://rdfs.org/sioc/ns#" }
  "@id": "",
  "@type": "sioc:Comment",
  "sioc:content": "This is a great article!",
  "sioc:reply_of": { "@id": "http://example.org/article" },
  "sioc:created_at": { "@value": "2015-12-23T16:44:21Z" }
}
\end{verbatim}

A notification about a comment created by a user (JSON-LD).

\begin{verbatim}
@prefix as: <https://www.w3.org/ns/activitystreams#> .
@prefix cito: <http://purl.org/spar/cito/> .
<> a as:Announce
  as:object <https://linkedresearch.org/resources#r-903b83> ;
  as:target <http://csarven.ca/dokieli#architecture> .
<https://linkedresearch.org/resources#r-903b83>
  cito:citesAsPotentialReading
    <http://csarven.ca/linked-data-notifications#protocol> .
\end{verbatim}

An announcement of a specific citation relation between two entities
(Turtle).

\hypertarget{implementations}{}
\hypertarget{implementations}{\subsection{Implementations}\label{implementations}}

Here we summarise the 17 LDN implementations we are aware of to date.
They are built by 10 different teams or individuals using different tool
stacks (5 clientside JavaScript, 3 PHP, 3 NodeJS, 3 Python, 1 Perl, 1
Virtuoso Server Pages, 1 Java) and have submitted
\href{https://github.com/w3c/ldn/tree/master/implementations}{implementation
reports} as part of the W3C standardisation process. We note that any
\href{https://www.w3.org/wiki/LDP_Implementations}{LDP implementation}
is a conforming LDN receiver; we refer here to the ones we have tested.
We discuss the value of these implementations further in the
\protect\hyperlink{analysis-and-evaluation}{Evaluation} section.

LDN Implementations

Implementation

Class\textsuperscript{\protect\hyperlink{implementations-key}{*}}

Description

\begin{description}
\tightlist
\item[\textsuperscript{*}]
Conformance classes: S -- sender, C -- consumer, R -- receiver.
\item[\textsuperscript{a}]
Implementations by the authors
\end{description}

Source: \url{https://github.com/w3c/ldn/tree/master/implementations}

\href{https://carbonldp.com}{CarbonLDP}

R

Data storage platform (LDP)

\href{https://dokie.li/}{dokieli}\textsuperscript{a}

S,C

Clientside editor and annotator

\href{https://github.com/linkeddata/errol}{errol}\textsuperscript{a}

S

Generic message sending client

\href{http://fedora-commons.org/}{Fedora Commons}

R

Open source repository platform (LDP)

\href{https://github.com/Kongaloosh/IndieAnndroid}{IndieAnndroid}

R

Personal blogging platform

\href{https://github.com/albertmeronyo/linked-edit-rules}{Linked Edit
Rules}

S

Statistical dataset consistency checker

\href{https://github.com/csarven/mayktso}{mayktso}\textsuperscript{a}

R

Personal data store (LDP)

\href{https://github.com/rhiaro/onscreen}{OnScreen}\textsuperscript{a}

C

Notifications display client

\href{https://github.com/albertmeronyo/pyldn}{pyldn}

R

Standalone Inbox

\href{https://github.com/kjetilk/p5-rdf-linkeddata-notifications}{RDF-LinkedData-Notifications}

R

Standalone Inbox

\href{https://github.com/rhiaro/sloph}{sloph}\textsuperscript{a}

S,R

Social publishing \& quantified self

\href{https://github.com/melvincarvalho/vocab}{Solid Words}

S

Foreign language learning app

\href{https://github.com/solid/solid-client}{solid-client}

S

Clientside library for LDP

\href{https://github.com/solid/solid-inbox}{solid-inbox}

C

Clientside social message reader

\href{https://github.com/solid/solid-notifications}{solid-notifications}

S,C

Clientside library for LDN

\href{https://github.com/solid/node-solid-server}{solid-server}

R

Personal data storage server (LDP)

\href{https://github.com/openlink/virtuoso-opensource}{Virtuoso}+\href{http://ods.openlinksw.com/wiki/ODS/OdsBriefcase}{ODS
Briefcase}

R,C

Personal data storage server (LDP)

We highlight social scholarly communication use cases with
\href{https://dokie.li/}{dokieli}, a clientside editor for decentralised
scientific article publishing, annotations and social interactions
{[}\protect\hyperlink{ref-17}{17}{]}. dokieli uses LDN to send and
consume notifications: When a reader comments on a fragment of text in
an article, the application discovers the article's Inbox and sends a
notification about the annotation. dokieli also consumes notifications
from this Inbox to fetch and display the annotation as marginalia
(\protect\hyperlink{figure-dokieli-annotation}{figure 2}). A reader can
share a dokieli-enabled article with their contacts; dokieli discovers
each contact's Inbox and sends a notification there
(\protect\hyperlink{figure-dokieli-share}{figure 3}). When editing an
article, the author can add a citation. If an Inbox is discovered in the
cited article, dokieli sends a notification there to indicate what part
of the article was cited by whom and where. dokieli-enabled articles
also consume citation notifications to display these metrics for the
author and other readers
(\protect\hyperlink{figure-dokieli-citation}{figure 4}).

\href{https://dokie.li/media/video/dokieli-annotation.webm}{Video} of
dokieli Web Annotation

\href{https://dokie.li/media/video/dokieli-share.webm}{Video} of dokieli
Share

Notifications sent by dokieli can be reused by any consuming
applications that recognise the vocabulary terms; similarly, dokieli can
consume notifications sent by different applications.

Further social use cases are demonstrated by
\href{https://rhiaro.co.uk/sloph}{sloph}, a personal publishing and
quantified self platform which acts as a node in a decentralised social
network. When new content is created on the server, sloph performs
discovery on URLs it finds as values of particular properties of the new
content, as well as any URLs in the body of the content, and sends
notifications accordingly. For instance:

\begin{itemize}
\tightlist
\item
  If a \emph{Like} activity is generated on the server, sloph uses the
  \texttt{object} of the \emph{Like} as the target for a notification.
  Since dokieli uses the same vocabulary for social interactions
  (\href{https://www.w3.org/ns/activitystreams-core}{ActivityStreams
  2.0} {[}\protect\hyperlink{ref-18}{18}{]}), if the target is a dokieli
  article, this \emph{Like} will be displayed
  (\protect\hyperlink{figure-sloph-dokieli}{figure 5}).
\item
  If the user publishes a blog post containing a link, which may be
  semantically annotated to indicate the reason for linking, sloph sends
  a notification to any Inbox discovered at that link.
\item
  As a receiver, sloph accepts all incoming notifications, but holds for
  moderation (i.e. places behind access control) any that it cannot
  automatically verify refer to third-party content published on another
  domain. If an article written with dokieli publishes a citation of a
  blog post which advertises a sloph Inbox, sloph will fetch the article
  and verify whether the relation matches the contents of the
  notification before exposing the notification for re-use.
\end{itemize}

\href{http://www.linkededitrules.org/}{Linked Edit Rules} and
\href{https://melvincarvalho.github.io/vocab/}{Solid Words} are
specialised senders. Linked Edit Rules checks the consistency of
statistical datasets against structured constraints, and delivers the
consistency report as a notification to the user. Solid Words is a
clientside game for learning new words in a foreign language; it
delivers the player's score for each round to their Inbox.
\href{https://apps.rhiaro.co.uk/onscreen}{OnScreen} is a (crude) generic
consumer; as such, it can display notifications sent by both of the
aforementioned senders (\protect\hyperlink{figure-ldn-senders}{figure
6}).

\includegraphics{media/images/screenshot-ldn-sloph-dokieli.jpg}

A \emph{Like} notification created by sloph, displayed by dokieli.

\includegraphics{media/images/screenshot-ldn-senders.jpg}

A: Solid Words (a sender), B: Linked Edit Rules (a sender), C: OnScreen
(a consumer) displaying notifications sent by A and B.

\hypertarget{analysis-and-evaluation}{}
\hypertarget{analysis-and-evaluation}{\subsection{Analysis and
Evaluation}\label{analysis-and-evaluation}}

The LDN protocol describes the discovery of a resource's Inbox whence
notifications are sent or consumed, and the sending and exposure of
those notifications. Here we analyse how well features of LDN achieve
the
\protect\hyperlink{requirements-and-design-considerations}{requirements}
identified previously, and compare this to related work.

We have already examined
\protect\hyperlink{implementations}{implementations} of the
specification and described how they interoperate with each other; this
can be further tested by running the
\href{https://linkedresearch.org/ldn/tests/}{test suite}:
https://linkedresearch.org/ldn/tests/. We can use this towards an
evaluation of its feasibility and effectiveness at interoperability.
Given the relatively early stage in the standardisation process (LDN
entered Candidate Recommendation in 2016-11), the fast adoption of the
LDN specification, quantity of the implementations, and their diversity
is promising and further shows LDN's feasibility. Furthermore, during
the development of the specification issues have been raised or
discussed by 28 different people (excluding the authors; 21 outside of
the Social Web Working Group, 7 within) and the specification has
undergone formal review by internationalisation, accessibility, and
security specialists. We also discuss in more depth particular
challenges that were raised and resolved as part of this process.

\hypertarget{comparison-summary}{}
\subsubsection{Comparison summary}\label{comparison-summary}

Here we compare existing notification mechanisms from related work. The
criteria includes our
\protect\hyperlink{requirements-and-design-considerations}{requirements
and design considerations} (\emph{Rx}) along with additional technical
information which helps to capture some design differences (\emph{Tx}).

Comparison of notification mechanisms

Mechanism

T1

T2

T3

R1

R2

R3

R4-A

R4-B

R4-C\textsuperscript{p}

R4-C\textsuperscript{v}

R4-C\textsuperscript{o}

R5

\begin{description}
\tightlist
\item[T1]
Notification type
\item[T2]
Delivery method
\item[T3]
Dependencies
\item[R1]
Modularity (application classes: S Sender, R Receiver, C Consumer, U
Subscriber/User)
\item[R2]
Reusability
\item[R3]
Persistence - required? how?
\item[R4-A]
Target representation
\item[R4-B]
Notification body
\item[R4-C\textsuperscript{p}]
Payload processing required?
\item[R4-C\textsuperscript{v}]
Verification - required? how?
\item[R4-C\textsuperscript{o}]
Requirements for referenced resources?
\item[R5]
Subscription
\end{description}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\begin{description}
\tightlist
\item[--]
not applicable, out of scope
\item[/]
not specified, in scope
\item[X]
explicitly disallowed
\item[app]
application specific decision
\item[!]
required (\emph{MUST})
\item[+]
recommended (\emph{SHOULD})
\item[O]
optional (\emph{MAY})
\item[PuSH]
PubSubHubbub
\end{description}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\begin{description}
\tightlist
\item[\textsuperscript{h}]
HTML recommended
\item[\textsuperscript{j}]
Alternate RDF formats can be negotiated
\item[\textsuperscript{k}]
\texttt{source} and \texttt{target} key--value pairs is required
\item[\textsuperscript{q}]
Provenance records with \href{http://www.w3.org/TR/prov-o/}{PROV
Ontology}
\item[\textsuperscript{r}]
RDF representation recommended
\item[\textsuperscript{ra}]
SPARQL results transformed to RSS/Atom
\item[\textsuperscript{s}]
\href{https://www.sitemaps.org/protocol.html}{Sitemaps}
\item[\textsuperscript{t}]
Described in an RDF store or dataset
\end{description}

Semantic Pingback

Linkback

POST

RDF

S R

/

/

Any\textsuperscript{r}

form urlencoded\textsuperscript{k}

!

! parse source

Any\textsuperscript{r}

X

Webmention

Linkback

POST

HTML

S R

--

--

Any\textsuperscript{h}

form urlencoded\textsuperscript{k}

!

! parse source

Any\textsuperscript{h}

X

Provenance Pingback

Linkback

POST

RDF

S R

/

/

/

URI list

/

/

RDF\textsuperscript{q}

X

DSNotify

Fat ping

POST, PUT

XML, PuSH

S U

/

--

--

XML

/

--

RDF\textsuperscript{t}

!

sparqlPuSH

Fat ping

POST

XML, SPARQL, PuSH

S U

--

--

--

XML\textsuperscript{ra}

/

--

RDF\textsuperscript{t}

!

ResourceSync

Fat ping

POST

XML, PuSH

S U

/

--

--

XML\textsuperscript{s}

/

--

?

!

Linked Data Notifications

Fat ping

POST

JSON-LD

S R C

!

! URI

Any

JSON-LD\textsuperscript{j}

+ app

+ app

--

O app

Given that each application requires to follow the steps listed in
``\protect\hyperlink{sender-to-receiver}{Sender to receiver
interaction}'' and ``\protect\hyperlink{consumer-to-receiver}{Consumer
to receiver interactions}'' the metrics are dependent on the performance
of client and server to do HTTP requests and responses, and their
respective payloads.

\hypertarget{compatibility-with-existing-systems}{}
\subsubsection{Compatibility with existing
systems}\label{compatibility-with-existing-systems}

Per \protect\hyperlink{modularity}{R1} and
\protect\hyperlink{adaptability}{R4} we have tried to optimise LDN for
use as a module of a larger system. The success of this is demonstrated
by implementations which use LDN alongside existing protocols according
to their specific needs.

The Solid suite of tools, Virtuoso+ODS-Briefcase, and dokieli use
\href{https://www.w3.org/wiki/WebAccessControl}{Web Access Control}
along with an authentication mechanism to apply fine grained access
controls to restrict who can send notifications, or who can retrieve
notifications from the Inbox. sloph demonstrates an Inbox as a
\href{http://www.webhooks.org/}{Webhooks} callback URL, for requesting
notifications from APIs which post JSON-based payloads.
\href{https://www.w3.org/TR/activitypub/}{ActivityPub} is a W3C CR for
decentralised social media {[}\protect\hyperlink{ref-19}{19}{]}. It uses
LDN for delivery of notifications with the
\href{https://www.w3.org/ns/activitystreams-vocabulary}{ActivityStreams
2.0} (AS2) vocabulary, and specifies additional specialised receiver
behaviour; also used by sloph. dokieli uses the
\href{https://www.w3.org/TR/annotation-protocol/}{Web Annotation
Protocol}, an LDP-based mechanism for creating new content, which acts
as a trigger for notifications to be sent to the Inbox of the annotation
target. The \href{http://fcrepo.github.io/fcrepo-specification/}{Fedora
API Specification} is in the process of being formalised (as an
extension of LDP) by the Fedora community. The repository event stream
draws upon the LDN specification, allowing LDN consumers and senders to
react asynchronously to repository events.

Any existing LDP implementation can serve as an LDN receiver. Simply
advertising any \texttt{ldp:Container} as the Inbox for a resource is
sufficient. We confirmed this with four LDP servers which were developed
independently with different code bases, prior to the LDN specification
(CarbonLDP, Fedora Commons, Solid Server, Virtuoso).

LDN has been integrated into existing domain specific systems: dokieli,
Fedora Commons, IndieAnndroid, Linked Edit Rules, sloph, solid-client,
Solid Words. Standalone implementations of LDN are also straightforward
as a result of this modularity, ie: errol, mayktso, onscreen, pyLDN,
RDF-LinkedData-Notifications, solid-inbox, solid-notifications.

\hypertarget{optimising-implementation}{}
\subsubsection{Optimising
implementation}\label{optimising-implementation}

{We have considered tradeoffs between the HTTP operations receivers and
publishers are \emph{required} to respond to, and ways in which
developers may wish to optimise senders or consumers by reducing
outbound requests.}

{{\texttt{HEAD} requests are low cost, and \texttt{GET} requests may be
high cost if the body of the resource is large.}} {{Given that an Inbox
may be discovered from the HTTP headers of a resource, senders and
consumers can optimise by attempting a \texttt{HEAD} request for
discovery, and only continuing with a \texttt{GET} request if the
\texttt{HEAD} is not successful. On the other hand, senders and
consumers may be attempting discovery upon RDF resources which they
already intend to parse into their own storage. In this case, there is
no need for a \texttt{HEAD} request, as a \texttt{GET} will yield both
HTTP \texttt{Link} headers and an RDF body, either of which could
include the Inbox triple. This means that resources advertising an Inbox
must respond to \texttt{GET} requests (even if only with HTTP headers)
and may respond to \texttt{HEAD} requests.}}

\hypertarget{data-formats}{}
\subsubsection{Data Formats and Content
Negotiation}\label{data-formats-and-content-negotiation}

{Handling data irrespective of the particular RDF serialisation permits
some flexibility, but can be costly to support. We take into account:
(a) application interoperability, (b) maintenance of RDF parsers and
serialisation libraries, (c) complexity of their inclusion in
applications, (d) run-time efficiency.}

{{To address these issues, LDN requires all applications to create and
understand the JSON-LD syntax, both for the contents of Inbox as well as
for individual notifications. Choosing a single serialisation to
\emph{require} is necessary for consistent interoperability, as well as
keeping processing requirements or external code dependencies minimal.}}
{{JSON-LD is advantageous in being familiar for developers who are
\href{http://manu.sporny.org/2014/json-ld-origins-2/}{used to JSON-based
APIs but not RDF} {[}\protect\hyperlink{ref-20}{20}{]}, and it is
compatible with existing JSON libraries or in some cases native
programming language data structures.}}

Optionally, {{applications may attempt to exchange different RDF
serialisations by performing content negotiation}} ({{receivers can
expose \texttt{Accept-Post} headers for senders, and consumers can send
\texttt{Accept} headers to receivers}}).

\hypertarget{precision}{}
\subsubsection{Precision}\label{precision}

In placing no constraints on the contained information, LDN enables a
sender to be precise and lossless with the data it is transmitting.
Approaches which send only URLs rely on the receiver interpreting a
third-party resource, which may or may not contain structured markup or
be under the control of the sender. Approaches which offer additional
guidance to aid the receiver in interpreting the source document(s)
nonetheless still restricts the sender. LDN therefore offers flexibility
to senders, increasing the potential uses for the notification
mechanism. LDN compensates for increased complexity on the receiver's
end by recommending filtering mechanisms, and moving some of the burden
of understanding notifications to the consumer role. As such LDN can
cover a broader variety of use cases.

\hypertarget{accommodating-different-targets}{}
\subsubsection{Accommodating different
targets}\label{accommodating-different-targets}

Per \emph{R4 Adaptability}, we want LDN to be available for all
resources in any publishing context. We consider lowering the bar for
publishers of target resources to be a worthwhile trade-off against
slightly increased complexity for senders and consumers. This is why we
require that senders and consumers must be equipped to discover Inboxes
through both HTTP headers and RDF content.

Since binary formats such as images and video cannot contain an RDF
relation, the HTTP header is essential for including them. It also
allows the inclusion of resources for which it is undesirable or
impractical to add individual Inbox relations, such as to elements in a
dataset; or circumstances where the developer responsible for the Inbox
relation is unable to modify the content. Conversely, non-informational
resources (represented with fragment URIs or 303 redirects) are unable
to express HTTP headers. Their relation to an Inbox must be expressed in
an RDF source. However, if a sender or consumer has a domain-specific
requirement to \emph{only} ever target non-informational resources, they
are exempt from the requirement of discovery via HTTP headers.

\hypertarget{conclusions}{}
\subsection{Conclusions}\label{conclusions}

In this article we describe LDN, a protocol for decentralised semantic
notifications, currently undergoing standardisation at the W3C. Key
elements are:

\begin{itemize}
\tightlist
\item
  Notifications as retrievable, reusable entities with their own URIs.
\item
  Distinct conformance classes for senders, receivers, and consumers.
\item
  Deliberately not defining the vocabulary of notification contents to
  allow for use in a range of different application domains.
\item
  Flexibility of authentication and verification, for the same reason.
\end{itemize}

We outlined design requirements, describe how LDN meets these, and
compare this with related work. We consider LDN to have greater
modularity and adaptability to different scenarios, as well as good
conformance with Linked Data principles. This specification has
potential to have high impact in increasing interoperability between
decentralised Linked Data applications in related domains, as well as
generating new discoverable content for the LOD Cloud. This is evidenced
by 17 diverse implementations which can be shown to interoperate with
each other, including generic libraries and datastores, and
domain-specific applications. Being on the W3C standards track increases
the likelihood of further adoption.

\hypertarget{references}{}
\subsection{References}\label{references}

\begin{enumerate}
\item
  \hypertarget{ref-1}{}

  Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem,
  A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid
  Platform for Social Web Applications, WWW, Demo, 2016,
  \url{http://www2016.net/proceedings/companion/p223.pdf}
\item
  \hypertarget{ref-2}{}

  Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An
  Architecture of a Distributed Semantic Social Network, Semantic Web
  Journal, 2012,
  \href{http://www.semantic-web-journal.net/sites/default/files/swj201_4.pdf}{}
\item
  \hypertarget{ref-3}{}

  Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N.,
  Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016,
  \url{http://events.linkeddata.org/ldow2016/papers/LDOW2016_paper_02.pdf}
\item
  \hypertarget{ref-4}{}

  Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate
  Recommendation, 2016, \url{https://www.w3.org/TR/ldn/}
\item
  \hypertarget{ref-5}{}

  Parecki, A.: Webmention, W3C Proposed Recommendation, 2016,
  \url{https://www.w3.org/TR/webmention/}
\item
  \hypertarget{ref-6}{}

  Langridge, S., Hickson, I.: Pingback 1.0, 2002,
  \url{http://www.hixie.ch/specs/pingback/pingback}
\item
  \hypertarget{ref-7}{}

  Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query, W3C Note,
  2013, \url{http://www.w3.org/TR/prov-aq/}
\item
  \hypertarget{ref-8}{}

  Haslhofer, B., Popitsch, N.: DSNotify -- Detecting and Fixing Broken
  Links in Linked Data Sets, WWW, 2010,
  \url{http://eprints.cs.univie.ac.at/81/1/2010_WWW_DSNotify.pdf}
\item
  \hypertarget{ref-9}{}

  Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data
  updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop
  Proceedings, Vol. 699, 2010,
  \url{http://ceur-ws.org/Vol-699/Paper6.pdf}
\item
  \hypertarget{ref-10}{}

  Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B.,
  Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework
  Specification -- Change Notification, 2016,
  \url{http://www.openarchives.org/rs/notification/1.0/notification}
\item
  \hypertarget{ref-11}{}

  Berners-Lee, T.: Cool URIs don't change, W3C, 1998,
  \url{https://www.w3.org/Provider/Style/URI.html}
\item
  \hypertarget{ref-12}{}

  Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On
  Persistent URIs, 2012,
  \url{http://philarcher.org/diary/2013/uripersistence/}
\item
  \hypertarget{ref-13}{}

  Fielding, R. T.: Architectural Styles and the Design of Network-based
  Software Architectures. Doctoral dissertation, University of
  California, Irvine, 2000,
  \url{http://www.ics.uci.edu/~fielding/pubs/dissertation/rest_arch_style.htm}
\item
  \hypertarget{ref-14}{}

  Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a
  match made for domain driven development?, WWW, WS-REST, 2011,
  \url{http://ws-rest.org/2011/proc/a5-page.pdf}
\item
  \hypertarget{ref-15}{}

  Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C
  Working Draft, 2016, \href{https://www.w3.org/TR/shacl/}{}
\item
  \hypertarget{ref-16}{}

  Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C
  Recommendation, 2015, \url{https://www.w3.org/TR/ldp/}
\item
  \hypertarget{ref-17}{}

  Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016,
  \url{http://csarven.ca/dokieli}
\item
  \hypertarget{ref-18}{}

  Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate
  Recommendation, 2016,
  \url{https://www.w3.org/TR/activitystreams-core/}
\item
  \hypertarget{ref-19}{}

  Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation,
  2016, \url{https://www.w3.org/TR/activitypub/}
\item
  \hypertarget{ref-20}{}

  Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014,
  \url{http://manu.sporny.org/2014/json-ld-origins-2/}
\end{enumerate}

\end{document}

### 2017-03-07 index.xh2latex.tex
abstract and thebibliography environments

--- 2017-03-07 index.tex
@@
 \documentclass[a4paper]{llncs}
-\usepackage{ifxetex}
-\ifxetex
-\usepackage{fontspec}
-\defaultfontfeatures{Ligatures=TeX}
-\else
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\fi
+\usepackage{lmodern}
+\usepackage[scaled=.8]{beramono}
+\usepackage{latexsym}
+\usepackage{textcomp}
 
 \usepackage{listings}
-\lstset{numbers=left, captionpos=b, frame=single, basicstyle=\ttfamily, stringstyle=\small, breaklines=true, extendedchars=true, showstringspaces=true, numberbychapter=false}
+%\lstset{numbers=left, captionpos=b, frame=single, basicstyle=\ttfamily, stringstyle=\small, breaklines=true, extendedchars=true, showstringspaces=true, numberbychapter=false}
 
-\usepackage{caption}
-\captionsetup{labelfont=bf}
-
-\usepackage{verbatim}
-
-\usepackage[numbers]{natbib}
-\usepackage{dcolumn}
-\usepackage[export]{adjustbox}
 \usepackage{graphicx}
-\usepackage[multidot]{grffile}
-\usepackage{url}
-\usepackage{xspace}
-\usepackage{color}
-\usepackage{float}
 \usepackage{multirow}
@@
 \usepackage{array}
-\usepackage{todonotes}
 \usepackage{hyperref}
@@
 
-                Linked Data Notifications: a resource-centric communication protocol
+\title{Linked Data Notifications: a resource-centric communication protocol}
 
-                    \begin{description}
-  \item[Authors] \empty Sarven Capadisli\empty 1\empty a\empty Amy Guy\empty 2\empty b\empty Christoph Lange\empty 1\empty c\empty Sören Auer\empty 1\empty d\empty Tim Berners-Lee\empty 3\empty e
-    \end{description}
+\author{Sarven Capadisli\inst{1}
+  \and Amy Guy\inst{2}
+  \and Christoph Lange\inst{1,3}
+  \and Sören Auer\inst{1,3}
+  \and Tim Berners-Lee\inst{4}}
+\institute{%
+  University of Bonn, Bonn, DE
+  \email{info@csarven.ca, \{langec,auer\}@cs.uni-bonn.de}
+  \and School of Informatics, University of Edinburgh, Edinburgh, UK
+  \email{amy@rhiaro.co.uk}
+  \and Fraunhofer IAIS, Sankt Augustin, DE
+  \and Decentralized Information Group, CSAIL, MIT, Cambridge, US
+  \email{timbl@w3.org}
+}
+\maketitle
 
-                    \begin{itemize}
-  \item 1Enterprise Information Systems Department, \empty University of Bonn, Bonn, DE\item 2School of Informatics, \empty University of Edinburgh, Edinburgh, UK\item 3Decentralized Information Group, CSAIL, \empty MIT, Cambridge, US
-    \end{itemize}
-
-                    \begin{itemize}
-  \item ✊\empty info@csarven.ca\item 🐦\empty amy@rhiaro.co.uk\item ∫\empty langec@cs.uni-bonn.de\item ⚛\empty auer@cs.uni-bonn.de\item 🕸\empty timbl@w3.org
-    \end{itemize}
-
-\par In this article we describe the Linked Data Notifications (LDN) protocol, which is a \empty W3C Candidate Recommendation.
+\begin{abstract}
+  In this article we describe the Linked Data Notifications (LDN) protocol, which is a W3C Candidate Recommendation.
 Notifications are sent over the Web for a variety of purposes, for example, by social applications.
@@
 To experience the described features please open this document in your Web browser under its canonical URI: {\tt \href{http://csarven.ca/linked-data-notifications}{http://csarven.ca/linked-data-notifications}}.
-
-                        \section{Keywords}
-  \label{keywords}
-
-                            \begin{itemize}
-  \item \empty Communications protocol\item \empty Decentralisation\item \empty Linked Data\item \empty Social web
-    \end{itemize}
+  \keywords{Communications protocol, Decentralisation, Linked Data, Social web}
+\end{abstract}
 
@@
 However in centralised systems which prevail today, this data is structured arbitrarily and typically only usable by the application that generated it in the first place.
-Current efforts towards {\em re-decentralising} the Web [\empty 1, \empty 2, \empty 3] are moving towards architectures in which data storage is decoupled from application logic, freeing end users to switch between applications, or to let multiple applications operate over the same data.
+Current efforts towards {\em re-decentralising} the Web~\cite{ref-1,ref-2,ref-3} are moving towards architectures in which data storage is decoupled from application logic, freeing end users to switch between applications, or to let multiple applications operate over the same data.
 So far, notifications are considered to be {\em ephemeral} resources which may disappear after transport, and thus are excluded from being designed for reuse.
@@
 
-\par LDN is a W3C \empty Candidate Recommendation via the \empty Social Web Working Group [\empty 4]. The first two authors of this article are the co-editors of the specification.
+\par LDN is a W3C \empty Candidate Recommendation via the \empty Social Web Working Group~\cite{ref-4}. The first two authors of this article are the co-editors of the specification.
 
@@
 
-\par \empty Semantic Pingback [\empty 2] and \empty Webmention [\empty 5] both update the original \empty Pingback [\empty 6] mechanism by replacing the XML-RPC transport mechanism by a {\tt x-www-form-urlencoded} request with two parameters ({\tt source} and {\tt target}).
+\par \empty Semantic Pingback~\cite{ref-2} and \empty Webmention~\cite{ref-5} both update the original \empty Pingback~\cite{ref-6} mechanism by replacing the XML-RPC transport mechanism by a {\tt x-www-form-urlencoded} request with two parameters ({\tt source} and {\tt target}).
 Resources which are the target for a notification advertise the respective receiving service or endpoint via a {\tt Link} relation, either in HTTP headers or HTML.
@@
 As such, there is also no way to distinguish between multiple potential mentions of the target at the source; this is left up to the receiver to interpret.
-Semantic Pingback does encourage generation of additional semantics about the relation(s) between the source and the target by processing the source as RDF if possible, and also defines specific ways for a receiving server to handle incoming pingback data in order to add the source data to an RDF knowledge base [\empty 2].
+Semantic Pingback does encourage generation of additional semantics about the relation(s) between the source and the target by processing the source as RDF if possible, and also defines specific ways for a receiving server to handle incoming pingback data in order to add the source data to an RDF knowledge base~\cite{ref-2}.
 Beyond verifying that the source contains the URL of the target, Webmention does not specify any further requirements of the receiving server; nor is it expected that “mentions” are retrievable once they have been sent.
 
-\par A \empty Provenance Pingback endpoint is also advertised via the HTTP {\tt Link} header; it accepts a list of URIs for provenance records describing uses of the resource [\empty 7].
+\par A \empty Provenance Pingback endpoint is also advertised via the HTTP {\tt Link} header; it accepts a list of URIs for provenance records describing uses of the resource~\cite{ref-7}.
 Provenance Pingback does not specify any further behaviour by the receiving server, but the contents at the URIs listed in the notification body must be semantic data.
@@
 \par \empty DSNotify is a centralised service which crawls datasets and observes changes to links with the specific use case of preserving link integrity between Linked Open Data resources.
-Third-party applications can register with the sending service to receive notifications of changes in the form of a specific XML payload [\empty 8].
+Third-party applications can register with the sending service to receive notifications of changes in the form of a specific XML payload~\cite{ref-8}.
 With the \empty sparqlPuSH service, users may input a SPARQL query, the results of which are the specific updates they are interested in.
-The query is run periodically by the service, and the results are converted to RSS and Atom feeds, which is sent to a \empty PubSubHubbub hub to which the user can subscribe [\empty 9].
-The \empty ResourceSync Change Notification specification also sends update notifications via a PuSH hub, this time with an XML payload based on the Sitemap format [\empty 10].
+The query is run periodically by the service, and the results are converted to RSS and Atom feeds, which is sent to a \empty PubSubHubbub hub to which the user can subscribe~\cite{ref-9}.
+The \empty ResourceSync Change Notification specification also sends update notifications via a PuSH hub, this time with an XML payload based on the Sitemap format~\cite{ref-10}.
 Each of these mechanisms is triggered by subscription requests.
@@
 
-\par There is a social expectation and technical arguments for ensuring the persistence of identifiers of Web resources [\empty 11].
+\par There is a social expectation and technical arguments for ensuring the persistence of identifiers of Web resources~\cite{ref-11}.
 This is inconsistent with the traditionally ephemeral nature of notifications.
-Applications may benefit from referring to or reusing notifications if the notifications are known to be available in the long term, or indicate their expected lifespan [\empty 12].
+Applications may benefit from referring to or reusing notifications if the notifications are known to be available in the long term, or indicate their expected lifespan~\cite{ref-12}.
 
-\par A {\em RESTful architecture} [\empty 13] is well suited for persistent notifications, as it involves organisation of atomic resources, their discovery and description, and a lightweight API for the CRUD (create, read, update, and delete) operations [\empty 14].
+\par A {\em RESTful architecture}~\cite{ref-13} is well suited for persistent notifications, as it involves organisation of atomic resources, their discovery and description, and a lightweight API for the CRUD (create, read, update, and delete) operations~\cite{ref-14}.
 This enforces the notion that notifications are considered resources in their own right, with their own dereferencable URIs.
@@
 If the notification constraints are also advertised by the receiving service as structured descriptions, generation and consumption of the notifications can be further automated.
-Possible specifications for doing so are W3C \empty Shapes Constraint Language (SHACL) [\empty 15] or \empty ShEx.
+Possible specifications for doing so are W3C \empty Shapes Constraint Language (SHACL)~\cite{ref-15} or \empty ShEx.
 
@@
 
-\par Given our adoption of Linked Data principles and a RESTful architecture, a further design decision was to ensure minimal compatibility with the \empty Linked Data Platform (LDP) specification [\empty 16].
+\par Given our adoption of Linked Data principles and a RESTful architecture, a further design decision was to ensure minimal compatibility with the \empty Linked Data Platform (LDP) specification~\cite{ref-16}.
 LDP is a RESTful read-write API for RDF resources, which groups related resources together into constructs known as ``Containers''.
@@
 It is not dependent on a complete implementation of LDP, but comprises an easy-to-implement subset.
-LDN is a \empty W3C Candidate Recommendation [\empty 4].
+LDN is a \empty W3C Candidate Recommendation~\cite{ref-4}.
 
@@
 
-\par We highlight social scholarly communication use cases with \empty dokieli, a clientside editor for decentralised scientific article publishing, annotations and social interactions [\empty 17].
+\par We highlight social scholarly communication use cases with \empty dokieli, a clientside editor for decentralised scientific article publishing, annotations and social interactions~\cite{ref-17}.
 dokieli uses LDN to send and consume notifications: When a reader comments on a fragment of text in an article, the application discovers the article’s Inbox and sends a notification about the annotation.
@@
   \item If a {\em Like} activity is generated on the server, sloph uses the {\tt object} of the {\em Like} as the target for a notification.
-Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0 [\empty 18]), if the target is a dokieli article, this {\em Like} will be displayed (\empty figure 5).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e.
+Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0~\cite{ref-18}), if the target is a dokieli article, this {\em Like} will be displayed (\empty figure 5).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e.
 places behind access control) any that it cannot automatically verify refer to third-party content published on another domain.
@@
 sloph demonstrates an Inbox as a \empty Webhooks callback URL, for requesting notifications from APIs which post JSON-based payloads.
-\empty ActivityPub is a W3C CR for decentralised social media [\empty 19].
+\empty ActivityPub is a W3C CR for decentralised social media~\cite{ref-19}.
 It uses LDN for delivery of notifications with the \empty ActivityStreams 2.0 (AS2) vocabulary, and specifies additional specialised receiver behaviour; also used by sloph.
@@
 
-                                        JSON-LD is advantageous in being familiar for developers who are \empty used to JSON-based APIs but not RDF [\empty 20], and it is compatible with existing JSON libraries or in some cases native programming language data structures.
+                                        JSON-LD is advantageous in being familiar for developers who are \empty used to JSON-based APIs but not RDF~\cite{ref-20}, and it is compatible with existing JSON libraries or in some cases native programming language data structures.
 
@@
 
-                            \begin{enumerate}
-  \item Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016, \empty http://www2016.net/proceedings/companion/p223.pdf\item Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012, \empty \item Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016, \empty http://events.linkeddata.org/ldow2016/papers/LDOW2016\_paper\_02.pdf\item Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/ldn/\item Parecki, A.: Webmention, W3C Proposed Recommendation, 2016, \empty https://www.w3.org/TR/webmention/\item Langridge, S., Hickson, I.: Pingback 1.0, 2002, \empty http://www.hixie.ch/specs/pingback/pingback\item Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \empty http://www.w3.org/TR/prov-aq/\item Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010, \empty http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf\item Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol.
-699, 2010, \empty http://ceur-ws.org/Vol-699/Paper6.pdf\item Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016, \empty http://www.openarchives.org/rs/notification/1.0/notification\item Berners-Lee, T.: Cool URIs don't change, W3C, 1998, \empty https://www.w3.org/Provider/Style/URI.html\item Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On Persistent URIs, 2012, \empty http://philarcher.org/diary/2013/uripersistence/\item Fielding, R.
+\begin{thebibliography}{99}
+  \bibitem{ref-1} Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016, \url{http://www2016.net/proceedings/companion/p223.pdf}
+  \bibitem{ref-2} Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012
+  \bibitem{ref-3} Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016, \url{http://events.linkeddata.org/ldow2016/papers/LDOW2016_paper_02.pdf}
+  \bibitem{ref-4} Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/ldn/}
+  \bibitem{ref-5} Parecki, A.: Webmention, W3C Proposed Recommendation, 2016,  \url{https://www.w3.org/TR/webmention/}
+  \bibitem{ref-6} Langridge, S., Hickson, I.: Pingback 1.0, 2002,  \url{http://www.hixie.ch/specs/pingback/pingback}
+  \bibitem{ref-7} Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \url{http://www.w3.org/TR/prov-aq/}
+  \bibitem{ref-8} Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010, \url{http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf}
+  \bibitem{ref-9} Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol.
+699, 2010, \url{http://ceur-ws.org/Vol-699/Paper6.pdf}
+  \bibitem{ref-10} Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016,  \url{http://www.openarchives.org/rs/notification/1.0/notification}
+  \bibitem{ref-11} Berners-Lee, T.: Cool URIs don't change, W3C, 1998,  \url{https://www.w3.org/Provider/Style/URI.html}
+  \bibitem{ref-12} Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On Persistent URIs, 2012, \url{http://philarcher.org/diary/2013/uripersistence/}
+  \bibitem{ref-13} Fielding, R.
 T.: Architectural Styles and the Design of Network-based Software Architectures.
-Doctoral dissertation, University of California, Irvine, 2000, \empty http://www.ics.uci.edu/\~fielding/pubs/dissertation/rest\_arch\_style.htm\item Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011, \empty http://ws-rest.org/2011/proc/a5-page.pdf\item Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016, \empty \item Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C Recommendation, 2015, \empty https://www.w3.org/TR/ldp/\item Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016, \empty http://csarven.ca/dokieli\item Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitystreams-core/\item Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitypub/\item Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014, \empty http://manu.sporny.org/2014/json-ld-origins-2/
-    \end{enumerate}
+Doctoral dissertation, University of California, Irvine, 2000,  \url{http://www.ics.uci.edu/~fielding/pubs/dissertation/rest_arch_style.htm}
+  \bibitem{ref-14} Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011,  \url{http://ws-rest.org/2011/proc/a5-page.pdf}
+  \bibitem{ref-15} Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016
+  \bibitem{ref-16} Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C Recommendation, 2015, \url{https://www.w3.org/TR/ldp/}
+  \bibitem{ref-17} Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016, \url{http://csarven.ca/dokieli}
+  \bibitem{ref-18} Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/activitystreams-core/}
+  \bibitem{ref-19} Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/activitypub/}
+  \bibitem{ref-20} Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014, \url{http://manu.sporny.org/2014/json-ld-origins-2/}
+\end{thebibliography}
 

### 2017-03-08 index.xh2latex.tex
using proper LaTeX; restored figures and tables

@@
 \usepackage{listings}
-%\lstset{numbers=left, captionpos=b, frame=single, basicstyle=\ttfamily, stringstyle=\small, breaklines=true, extendedchars=true, showstringspaces=true, numberbychapter=false}
+\lstset{numbers=left, captionpos=b, frame=single, basicstyle=\ttfamily, stringstyle=\small, breaklines=true, extendedchars=true, showstringspaces=true, numberbychapter=false}
 
@@
 \usepackage{tabularx}
+\usepackage{ctable}
 \usepackage{array}
@@
 It is not dependent on a complete implementation of LDP, but comprises an easy-to-implement subset.
-LDN is a \empty W3C Candidate Recommendation~\cite{ref-4}.
 
-                                Overview of Linked Data Notifications
+\begin{figure}
+  \centering
+  \includegraphics[trim=.2cm 10cm .5cm 10cm,clip,width=.6\textwidth]{media/images/{linked-data-notifications-overview.svg}.pdf}
+  \caption{Overview of Linked Data Notifications}
+  \label{fig:overview}
+\end{figure}
 
@@
 
-                                        \begin{verbatim}{
-  "@context": { "sioc": "http://rdfs.org/sioc/ns#" }
+\begin{lstlisting}
+{ "@context": { "sioc": "http://rdfs.org/sioc/ns#" }
   "@id": "",
@@
   "sioc:reply_of": { "@id": "http://example.org/article" },
-  "sioc:created_at": { "@value": "2015-12-23T16:44:21Z" }
-}\end{verbatim}
+  "sioc:created_at": { "@value": "2015-12-23T16:44:21Z" } }
+\end{lstlisting}
                                         A notification about a comment created by a user (JSON-LD).
 
-                                        \begin{verbatim}@prefix as: <https://www.w3.org/ns/activitystreams#> .
+\begin{lstlisting}
+@prefix as: <https://www.w3.org/ns/activitystreams#> .
 @prefix cito: <http://purl.org/spar/cito/> .
@@
   cito:citesAsPotentialReading
-    <http://csarven.ca/linked-data-notifications#protocol> .\end{verbatim}
+    <http://csarven.ca/linked-data-notifications#protocol> .
+\end{lstlisting}
                                         An announcement of a specific citation relation between two entities (Turtle).
@@
 
-                                LDN Implementations
+\ctable[
+  caption={LDN Implementations},
+  label = {tab:implementations}
+]{lll}{
+  \tnote[*]{Conformance classes: S – sender, C – consumer, R – receiver.}
+  \tnote[a]{Implementations by the authors}
+  \tnote[]{Source: \url{https://github.com/w3c/ldn/tree/master/implementations}}
+}{\FL
+  Implementation & Class\tmark[*] & Description
+  \ML
+                                        CarbonLDP &
+                                        R &
+                                        Data storage platform (LDP) \NN
 
-                                        Implementation
-                                        Class\empty *
-                                        Description
+                                        dokieli\tmark[a] &
+                                        S,C &
+                                        Clientside editor and annotator \NN
 
-                                        \begin{description}
-  \item[*] Conformance classes: S – sender, C – consumer, R – receiver.\item[a] Implementations by the authors
-    \end{description}
+                                        errol\tmark[a] &
+                                        S &
+                                        Generic message sending client \NN
 
-\par Source: \empty https://github.com/w3c/ldn/tree/master/implementations
+                                        Fedora Commons &
+                                        R &
+                                        Open source repository platform (LDP) \NN
 
-                                        \empty CarbonLDP
-                                        R
-                                        Data storage platform (LDP)
+                                        IndieAnndroid &
+                                        R &
+                                        Personal blogging platform \NN
 
-                                        \empty dokielia
-                                        S,C
-                                        Clientside editor and annotator
+                                        Linked Edit Rules &
+                                        S &
+                                        Statistical dataset consistency checker \NN
 
-                                        \empty errola
-                                        S
-                                        Generic message sending client
+                                        mayktso\tmark[a] &
+                                        R &
+                                        Personal data store (LDP) \NN
 
-                                        \empty Fedora Commons
-                                        R
-                                        Open source repository platform (LDP)
+                                        OnScreen\tmark[a] &
+                                        C &
+                                        Notifications display client \NN
 
-                                        \empty IndieAnndroid
-                                        R
-                                        Personal blogging platform
+                                        pyldn &
+                                        R &
+                                        Standalone Inbox \NN
 
-                                        \empty Linked Edit Rules
-                                        S
-                                        Statistical dataset consistency checker
+                                        RDF-LinkedData-Notifications &
+                                        R &
+                                        Standalone Inbox \NN
 
-                                        \empty mayktsoa
-                                        R
-                                        Personal data store (LDP)
+                                        sloph\tmark[a] &
+                                        S,R &
+                                        Social publishing \& quantified self \NN
 
-                                        \empty OnScreena
-                                        C
-                                        Notifications display client
+                                        Solid Words &
+                                        S &
+                                        Foreign language learning app \NN
 
-                                        \empty pyldn
-                                        R
-                                        Standalone Inbox
+                                        solid-client &
+                                        S &
+                                        Clientside library for LDP \NN
 
-                                        \empty RDF-LinkedData-Notifications
-                                        R
-                                        Standalone Inbox
+                                        solid-inbox &
+                                        C &
+                                        Clientside social message reader \NN
 
-                                        \empty slopha
-                                        S,R
-                                        Social publishing \& quantified self
+                                        solid-notifications &
+                                        S,C &
+                                        Clientside library for LDN \NN
 
-                                        \empty Solid Words
-                                        S
-                                        Foreign language learning app
+                                        solid-server &
+                                        R &
+                                        Personal data storage server (LDP) \NN
 
-                                        \empty solid-client
-                                        S
-                                        Clientside library for LDP
-
-                                        \empty solid-inbox
-                                        C
-                                        Clientside social message reader
-
-                                        \empty solid-notifications
-                                        S,C
-                                        Clientside library for LDN
-
-                                        \empty solid-server
-                                        R
-                                        Personal data storage server (LDP)
-
-                                        \empty Virtuoso+\empty ODS Briefcase
-                                        R,C
-                                        Personal data storage server (LDP)
+                                        Virtuoso+ ODS Briefcase &
+                                        R,C &
+                                        Personal data storage server (LDP) \LL
+}
 
@@
 dokieli uses LDN to send and consume notifications: When a reader comments on a fragment of text in an article, the application discovers the article’s Inbox and sends a notification about the annotation.
-dokieli also consumes notifications from this Inbox to fetch and display the annotation as marginalia (\empty figure 2).
-A reader can share a dokieli-enabled article with their contacts; dokieli discovers each contact’s Inbox and sends a notification there (\empty figure 3).
+dokieli also consumes notifications from this Inbox to fetch and display the annotation as marginalia (figure~\ref{fig:video-annotation}).
+A reader can share a dokieli-enabled article with their contacts; dokieli discovers each contact’s Inbox and sends a notification there (figure~\ref{fig:video-share}).
 When editing an article, the author can add a citation.
 If an Inbox is discovered in the cited article, dokieli sends a notification there to indicate what part of the article was cited by whom and where.
-dokieli-enabled articles also consume citation notifications to display these metrics for the author and other readers (\empty figure 4).
+dokieli-enabled articles also consume citation notifications to display these metrics for the author and other readers (figure~\ref{fig:like}).
 
-                                \empty Video of dokieli Web Annotation
-
-                                \empty Video of dokieli Share
+\begin{figure}
+  \begin{minipage}[b]{.49\textwidth}
+    \includegraphics[width=\textwidth]{media/images/dokieli-annotation}
+    \caption{Video of dokieli Web Annotation}
+    \label{fig:video-annotation}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{.49\textwidth}
+    \includegraphics[width=\textwidth]{media/images/dokieli-share}
+    \caption{Video of dokieli Share}
+    \label{fig:video-share}
+  \end{minipage}
+\end{figure}
 
@@
   \item If a {\em Like} activity is generated on the server, sloph uses the {\tt object} of the {\em Like} as the target for a notification.
-Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0~\cite{ref-18}), if the target is a dokieli article, this {\em Like} will be displayed (\empty figure 5).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e.
+Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0~\cite{ref-18}), if the target is a dokieli article, this {\em Like} will be displayed (figure~\ref{fig:like}).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e.
 places behind access control) any that it cannot automatically verify refer to third-party content published on another domain.
@@
 Solid Words is a clientside game for learning new words in a foreign language; it delivers the player’s score for each round to their Inbox.
-\empty OnScreen is a (crude) generic consumer; as such, it can display notifications sent by both of the aforementioned senders (\empty figure 6).
+\empty OnScreen is a (crude) generic consumer; as such, it can display notifications sent by both of the aforementioned senders (figure~\ref{fig:notifications}).
 
-                                A {\em Like} notification created by sloph, displayed by dokieli.
+\begin{figure}
+  \includegraphics[width=\textwidth]{media/images/screenshot-ldn-sloph-dokieli}
+  \caption{A {\em Like} notification created by sloph, displayed by dokieli.}
+  \label{fig:like}
 
-                                A: Solid Words (a sender), B: Linked Edit Rules (a sender), C: OnScreen (a consumer) displaying notifications sent by A and B.
+  \includegraphics[width=\textwidth]{media/images/screenshot-ldn-senders}
+  \caption{A: Solid Words (a sender), B: Linked Edit Rules (a sender), C: OnScreen (a consumer) displaying notifications sent by A and B.}
+  \label{fig:notifications}
+\end{figure}
 
@@
 
-                                        Comparison of notification mechanisms
-
-                                                Mechanism
-                                                T1
-                                                T2
-                                                T3
-                                                R1
-                                                R2
-                                                R3
-                                                R4-A
-                                                R4-B
-                                                R4-Cp
-                                                R4-Cv
-                                                R4-Co
-                                                R5
-
-                                                    \begin{description}
-  \item[T1] Notification type\item[T2] Delivery method\item[T3] Dependencies\item[R1] Modularity (application classes: S Sender, R Receiver, C Consumer, U Subscriber/User)\item[R2] Reusability\item[R3] Persistence - required?
-how?\item[R4-A] Target representation\item[R4-B] Notification body\item[R4-Cp] Payload processing required?\item[R4-Cv] Verification - required?
-how?\item[R4-Co] Requirements for referenced resources?\item[R5] Subscription
-    \end{description}
-
-                                                    \begin{description}
-  \item[–] not applicable, out of scope\item[/] not specified, in scope\item[X] explicitly disallowed\item[app] application specific decision\item[!] required ({\em MUST})\item[+] recommended ({\em SHOULD})\item[O] optional ({\em MAY})\item[PuSH] PubSubHubbub
-    \end{description}
-
-                                                    \begin{description}
-  \item[h] HTML recommended\item[j] Alternate RDF formats can be negotiated\item[k] {\tt source} and {\tt target} key–value pairs is required\item[q] Provenance records with \empty PROV Ontology\item[r] RDF representation recommended\item[ra] SPARQL results transformed to RSS/Atom\item[s] \empty Sitemaps\item[t] Described in an RDF store or dataset
-    \end{description}
-
-                                                Semantic Pingback
-                                                Linkback
-                                                POST
-                                                RDF
-                                                S R
-                                                /
-                                                /
-                                                Anyr
-                                                form urlencodedk
-                                                !
-                                                ! parse source
-                                                Anyr
-                                                X
-
-                                                Webmention
-                                                Linkback
-                                                POST
-                                                HTML
-                                                S R
-                                                –
-                                                –
-                                                Anyh
-                                                form urlencodedk
-                                                !
-                                                ! parse source
-                                                Anyh
-                                                X
-
-                                                Provenance Pingback
-                                                Linkback
-                                                POST
-                                                RDF
-                                                S R
-                                                /
-                                                /
-                                                /
-                                                URI list
-                                                /
-                                                /
-                                                RDFq
-                                                X
-
-                                                DSNotify
-                                                Fat ping
-                                                POST, PUT
-                                                XML, PuSH
-                                                S U
-                                                /
-                                                –
-                                                –
-                                                XML
-                                                /
-                                                –
-                                                RDFt
-                                                !
-
-                                                sparqlPuSH
-                                                Fat ping
-                                                POST
-                                                XML, SPARQL, PuSH
-                                                S U
-                                                –
-                                                –
-                                                –
-                                                XMLra
-                                                /
-                                                –
-                                                RDFt
-                                                !
-
-                                                ResourceSync
-                                                Fat ping
-                                                POST
-                                                XML, PuSH
-                                                S U
-                                                /
-                                                –
-                                                –
-                                                XMLs
-                                                /
-                                                –
-                                                ?
-                                                !
-
-                                                Linked Data Notifications
-                                                Fat ping
-                                                POST
-                                                JSON-LD
-                                                S R C
-                                                !
-                                                ! URI
-                                                Any
-                                                JSON-LDj
-                                                + app
-                                                + app
-                                                –
-                                                O app
+\begin{table}
+  \caption{Comparison of notification mechanisms}
+  \label{tab:comparison}
+  \footnotesize
+  \begin{tabular}{% Mechanism
+    p{.15\textwidth}
+    % T1–T3
+    p{.07\textwidth}
+    p{.07\textwidth}
+    p{.09\textwidth}
+    % R1–R3
+    p{.03\textwidth}
+    p{.03\textwidth}
+    p{.05\textwidth}
+    % R4-A–R5
+    p{.06\textwidth}
+    p{.14\textwidth}
+    p{.045\textwidth}
+    p{.06\textwidth}
+    p{.06\textwidth}
+    p{.05\textwidth}
+    }\FL
+                                                Mechanism&
+                                                T1&
+                                                T2&
+                                                T3&
+                                                R1&
+                                                R2&
+                                                R3&
+                                                R4-A&
+                                                R4-B&
+                                                R4-C\textsuperscript{p}&
+                                                R4-C\textsuperscript{v}&
+                                                R4-C\textsuperscript{o}&
+                                                R5\ML
+                                                Semantic\newline Pingback&
+                                                Link-\newline back&
+                                                POST&
+                                                RDF&
+                                                S\newline R&
+                                                /&
+                                                /&
+                                                Any\textsuperscript{r}&
+                                                form\newline urlencoded\textsuperscript{k}&
+                                                !&
+                                                !\newline parse\newline source&
+                                                Any\textsuperscript{r}&
+                                                X\NN
+                                                Webmention&
+                                                Link-\newline back&
+                                                POST&
+                                                HTML&
+                                                S\newline R&
+                                                –&
+                                                –&
+                                                Any\textsuperscript{h}&
+                                                form\newline urlencoded\textsuperscript{k}&
+                                                !&
+                                                !\newline parse\newline source&
+                                                Any\textsuperscript{h}&
+                                                X\NN
+                                                Provenance\newline Pingback&
+                                                Link-\newline back&
+                                                POST&
+                                                RDF&
+                                                S\newline R&
+                                                /&
+                                                /&
+                                                /&
+                                                URI list&
+                                                /&
+                                                /&
+                                                RDF\textsuperscript{q}&
+                                                X\NN
+                                                DSNotify&
+                                                Fat\newline ping&
+                                                POST,\newline PUT&
+                                                XML,\newline PuSH&
+                                                S\newline U&
+                                                /&
+                                                –&
+                                                –&
+                                                XML&
+                                                /&
+                                                –&
+                                                RDF\textsuperscript{t}&
+                                                !\NN
+                                                sparqlPuSH&
+                                                Fat\newline ping&
+                                                POST&
+                                                XML,\newline SPARQL,\newline PuSH&
+                                                S\newline U&
+                                                –&
+                                                –&
+                                                –&
+                                                XML\textsuperscript{ra}&
+                                                /&
+                                                –&
+                                                RDF\textsuperscript{t}&
+                                                !\NN
+                                                ResourceSync&
+                                                Fat\newline ping&
+                                                POST&
+                                                XML,\newline PuSH&
+                                                S\newline U&
+                                                /&
+                                                –&
+                                                –&
+                                                XML\textsuperscript{s}&
+                                                /&
+                                                –&
+                                                ?&
+                                                !\NN
+                                                Linked Data\newline Notifications&
+                                                Fat\newline ping&
+                                                POST&
+                                                JSON-LD&
+                                                S\newline R\newline C&
+                                                !&
+                                                !\newline URI&
+                                                Any&
+                                                JSON-LD\textsuperscript{j}&
+                                                + app&
+                                                + app&
+                                                –&
+                                                O\newline app\LL
+  \end{tabular}\\
+  \begin{tabularx}{1.0\linewidth}{XX}\FL
+    \textbf{T1}: Notification type&
+    \textbf{R4-A}: Target representation\NN
+    \textbf{T2}: Delivery method&
+    \textbf{R4-B}: Notification body\NN
+    \textbf{T3}: Dependencies&
+    \textbf{R4-C\textsuperscript{p}}: Payload processing required?\NN
+    \textbf{R1}: Modularity (application classes: S Sender, R Receiver, C Consumer, U Subscriber/User)&
+    \textbf{R4-C\textsuperscript{v}}: Verification - required? how?\NN
+    \textbf{R2}: Reusability&
+    \textbf{R4-C\textsuperscript{o}}: Requirements for referenced resources?\NN
+    \textbf{R3}: Persistence - required? how?&
+    \textbf{R5}: Subscription\ML
+    \textbf{–}: not applicable, out of scope&
+    \textbf{!}: required ({\em MUST})\NN
+    \textbf{/}: not specified, in scope&
+    \textbf{+}: recommended ({\em SHOULD})\NN
+    \textbf{X}: explicitly disallowed&
+    \textbf{O}: optional ({\em MAY})\NN
+    \textbf{app}: application specific decision&
+    \textbf{PuSH}: PubSubHubbub\ML
+    \textbf{\textsuperscript{h}}: HTML recommended&
+    \textbf{\textsuperscript{r}}: RDF representation recommended\NN
+    \textbf{\textsuperscript{j}}: Alternate RDF formats can be negotiated&
+    \textbf{\textsuperscript{ra}}: SPARQL results transformed to RSS/Atom\NN
+    \textbf{\textsuperscript{k}}: {\tt source} and {\tt target} key–value pairs is required&
+    \textbf{\textsuperscript{s}}: \empty Sitemaps\NN
+    \textbf{\textsuperscript{q}}: Provenance records with \empty PROV Ontology&
+    \textbf{\textsuperscript{t}}: Described in an RDF store or dataset\LL
+  \end{tabularx}
+\end{table}
 
@@
 
+\paragraph{Acknowledgements:}
+This research was supported in part by Massachusetts Institute of Technology and Qatar Computing Research Institute, HBKU through the Crosscloud project from 2015-10 to 2016-09, and by the DFG project “Opening Scholarly Communication in Social Sciences” (grant agreement AU 340/9-1).
+
 \begin{thebibliography}{99}

### 2017-03-08 index.xh2latex.tex
fine-tuning; commenting some bib URLs

@@
   \begin{tabular}{% Mechanism
-    p{.15\textwidth}
+    p{.14\textwidth}
     % T1–T3
-    p{.07\textwidth}
-    p{.07\textwidth}
-    p{.09\textwidth}
+    p{.065\textwidth}
+    p{.075\textwidth}
+    p{.1\textwidth}
     % R1–R3
@@
     p{.03\textwidth}
-    p{.05\textwidth}
+    p{.04\textwidth}
     % R4-A–R5
@@
                                                 X\NN
-                                                Webmention&
+                                                Web-\newline mention&
                                                 Link-\newline back&
@@
                                                 !\NN
-                                                ResourceSync&
+                                                Resource-\newline Sync&
                                                 Fat\newline ping&
@@
 \begin{thebibliography}{99}
-  \bibitem{ref-1} Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016, \url{http://www2016.net/proceedings/companion/p223.pdf}
+  \bibitem{ref-1} Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016%, \url{http://www2016.net/proceedings/companion/p223.pdf}
   \bibitem{ref-2} Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012
-  \bibitem{ref-3} Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016, \url{http://events.linkeddata.org/ldow2016/papers/LDOW2016_paper_02.pdf}
+  \bibitem{ref-3} Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016%, \url{http://events.linkeddata.org/ldow2016/papers/LDOW2016_paper_02.pdf}
   \bibitem{ref-4} Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/ldn/}
@@
   \bibitem{ref-7} Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \url{http://www.w3.org/TR/prov-aq/}
-  \bibitem{ref-8} Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010, \url{http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf}
+  \bibitem{ref-8} Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010%, \url{http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf}
   \bibitem{ref-9} Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol.
-699, 2010, \url{http://ceur-ws.org/Vol-699/Paper6.pdf}
+699, 2010%, \url{http://ceur-ws.org/Vol-699/Paper6.pdf}
   \bibitem{ref-10} Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016,  \url{http://www.openarchives.org/rs/notification/1.0/notification}
@@
 T.: Architectural Styles and the Design of Network-based Software Architectures.
-Doctoral dissertation, University of California, Irvine, 2000,  \url{http://www.ics.uci.edu/~fielding/pubs/dissertation/rest_arch_style.htm}
-  \bibitem{ref-14} Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011,  \url{http://ws-rest.org/2011/proc/a5-page.pdf}
+Doctoral dissertation, University of California, Irvine, 2000%,  \url{http://www.ics.uci.edu/~fielding/pubs/dissertation/rest_arch_style.htm}
+  \bibitem{ref-14} Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011%,  \url{http://ws-rest.org/2011/proc/a5-page.pdf}
   \bibitem{ref-15} Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016

### 2017-03-17 index.tex
Add Andrei and update acknowledgements

@@
 \documentclass[a4paper]{llncs}
-\usepackage{ifxetex}
-\ifxetex
-\usepackage{fontspec}
-\defaultfontfeatures{Ligatures=TeX}
-\else
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\fi
+\usepackage{lmodern}
+\usepackage[scaled=.8]{beramono}
+\usepackage{latexsym}
+\usepackage{textcomp}
 
@@
 
-\usepackage{caption}
-\captionsetup{labelfont=bf}
-
-\usepackage{verbatim}
-
-\usepackage[numbers]{natbib}
-\usepackage{dcolumn}
-\usepackage[export]{adjustbox}
 \usepackage{graphicx}
-\usepackage[multidot]{grffile}
-\usepackage{url}
-\usepackage{xspace}
-\usepackage{color}
-\usepackage{float}
 \usepackage{multirow}
@@
 \usepackage{tabularx}
+\usepackage{ctable}
 \usepackage{array}
-\usepackage{todonotes}
 \usepackage{hyperref}
@@
 
-                Linked Data Notifications: a resource-centric communication protocol
+\title{Linked Data Notifications: a resource-centric communication protocol}
 
-                    \begin{description}
-  \item[Authors] \empty Sarven Capadisli\empty 1\empty a\empty Amy Guy\empty 2\empty b\empty Christoph Lange\empty 1\empty c\empty Sören Auer\empty 1\empty d\empty Tim Berners-Lee\empty 3\empty e
-    \end{description}
+\author{Sarven Capadisli\inst{1}
+  \and Amy Guy\inst{2}
+  \and Christoph Lange\inst{1,3}
+  \and Sören Auer\inst{1,3}
+  \and Andrei Sambra\inst{4}
+  \and Tim Berners-Lee\inst{4}}
+\institute{%
+  University of Bonn, Bonn, DE
+  \email{info@csarven.ca, \{langec,auer\}@cs.uni-bonn.de}
+  \and School of Informatics, University of Edinburgh, Edinburgh, UK
+  \email{amy@rhiaro.co.uk}
+  \and Fraunhofer IAIS, Sankt Augustin, DE
+  \and Decentralized Information Group, CSAIL, MIT, Cambridge, US
+  \email{deiu@mit.edu, timbl@w3.org}
+}
+\maketitle
 
-                    \begin{itemize}
-  \item 1Enterprise Information Systems Department, \empty University of Bonn, Bonn, DE\item 2School of Informatics, \empty University of Edinburgh, Edinburgh, UK\item 3Decentralized Information Group, CSAIL, \empty MIT, Cambridge, US
-    \end{itemize}
-
-                    \begin{itemize}
-  \item ✊\empty info@csarven.ca\item 🐦\empty amy@rhiaro.co.uk\item ∫\empty langec@cs.uni-bonn.de\item ⚛\empty auer@cs.uni-bonn.de\item 🕸\empty timbl@w3.org
-    \end{itemize}
-
-\par In this article we describe the Linked Data Notifications (LDN) protocol, which is a \empty W3C Candidate Recommendation.
+\begin{abstract}
+  In this article we describe the Linked Data Notifications (LDN) protocol, which is a W3C Candidate Recommendation.
 Notifications are sent over the Web for a variety of purposes, for example, by social applications.
@@
 It is not dependent on a complete implementation of LDP, but comprises an easy-to-implement subset.
-LDN is a \empty W3C Candidate Recommendation [\empty 4].
 
-                                Overview of Linked Data Notifications
+\begin{figure}
+  \centering
+  \includegraphics[trim=.2cm 10cm .5cm 10cm,clip,width=.6\textwidth]{media/images/{linked-data-notifications-overview.svg}.pdf}
+  \caption{Overview of Linked Data Notifications}
+  \label{fig:overview}
+\end{figure}
 
@@
 
-                                LDN Implementations
+\ctable[
+  caption={LDN Implementations},
+  label = {tab:implementations}
+]{lll}{
+  \tnote[*]{Conformance classes: S – sender, C – consumer, R – receiver.}
+  \tnote[a]{Implementations by the authors}
+  \tnote[]{Source: \url{https://github.com/w3c/ldn}}
+}{\FL
+  Implementation & Class\tmark[*] & Description
+  \ML
+                                        CarbonLDP &
+                                        R &
+                                        Data storage platform (LDP) \NN
 
-                                        Implementation
-                                        Class\empty *
-                                        Description
+                                        dokieli\tmark[a] &
+                                        S,C &
+                                        Clientside editor and annotator \NN
 
-                                        \begin{description}
-  \item[*] Conformance classes: S – sender, C – consumer, R – receiver.\item[a] Implementations by the authors
-    \end{description}
+                                        errol\tmark[a] &
+                                        S &
+                                        Generic message sending client \NN
 
-\par Source: \empty https://github.com/w3c/ldn/tree/master/implementations
+                                        Fedora Commons &
+                                        R &
+                                        Open source repository platform (LDP) \NN
 
-                                        \empty CarbonLDP
-                                        R
-                                        Data storage platform (LDP)
+                                        IndieAnndroid &
+                                        R &
+                                        Personal blogging platform \NN
 
-                                        \empty dokielia
-                                        S,C
-                                        Clientside editor and annotator
+                                        Linked Edit Rules &
+                                        S &
+                                        Statistical dataset consistency checker \NN
 
-                                        \empty errola
-                                        S
-                                        Generic message sending client
+                                        mayktso\tmark[a] &
+                                        R &
+                                        Personal data store (LDP) \NN
 
-                                        \empty Fedora Commons
-                                        R
-                                        Open source repository platform (LDP)
+                                        OnScreen\tmark[a] &
+                                        C &
+                                        Notifications display client \NN
 
-                                        \empty IndieAnndroid
-                                        R
-                                        Personal blogging platform
+                                        pyldn &
+                                        R &
+                                        Standalone Inbox \NN
 
-                                        \empty Linked Edit Rules
-                                        S
-                                        Statistical dataset consistency checker
+                                        RDF-LinkedData-Notifications &
+                                        R &
+                                        Standalone Inbox \NN
 
-                                        \empty mayktsoa
-                                        R
-                                        Personal data store (LDP)
+                                        sloph\tmark[a] &
+                                        S,R &
+                                        Social publishing \& quantified self \NN
 
-                                        \empty OnScreena
-                                        C
-                                        Notifications display client
+                                        Solid Words &
+                                        S &
+                                        Foreign language learning app \NN
 
-                                        \empty pyldn
-                                        R
-                                        Standalone Inbox
+                                        solid-client &
+                                        S &
+                                        Clientside library for LDP \NN
 
-                                        \empty RDF-LinkedData-Notifications
-                                        R
-                                        Standalone Inbox
+                                        solid-inbox &
+                                        C &
+                                        Clientside social message reader \NN
 
-                                        \empty slopha
-                                        S,R
-                                        Social publishing \& quantified self
+                                        solid-notifications &
+                                        S,C &
+                                        Clientside library for LDN \NN
 
-                                        \empty Solid Words
-                                        S
-                                        Foreign language learning app
+                                        solid-server &
+                                        R &
+                                        Personal data storage server (LDP) \NN
 
-                                        \empty solid-client
-                                        S
-                                        Clientside library for LDP
+                                        Virtuoso+ ODS Briefcase &
+                                        R,C &
+                                        Personal data storage server (LDP) \LL
+}
 
-                                        \empty solid-inbox
-                                        C
-                                        Clientside social message reader
-
-                                        \empty solid-notifications
-                                        S,C
-                                        Clientside library for LDN
-
-                                        \empty solid-server
-                                        R
-                                        Personal data storage server (LDP)
-
-                                        \empty Virtuoso+\empty ODS Briefcase
-                                        R,C
-                                        Personal data storage server (LDP)
-
-\par We highlight social scholarly communication use cases with \empty dokieli, a clientside editor for decentralised scientific article publishing, annotations and social interactions [\empty 17].
+\par We highlight social scholarly communication use cases with \empty dokieli, a clientside editor for decentralised scientific article publishing, annotations and social interactions~\cite{ref-17}.
 dokieli uses LDN to send and consume notifications: When a reader comments on a fragment of text in an article, the application discovers the article’s Inbox and sends a notification about the annotation.
-dokieli also consumes notifications from this Inbox to fetch and display the annotation as marginalia (\empty figure 2).
-A reader can share a dokieli-enabled article with their contacts; dokieli discovers each contact’s Inbox and sends a notification there (\empty figure 3).
+dokieli also consumes notifications from this Inbox to fetch and display the annotation as marginalia (figure~\ref{fig:video-annotation}).
+A reader can share a dokieli-enabled article with their contacts; dokieli discovers each contact’s Inbox and sends a notification there (figure~\ref{fig:video-share}).
 When editing an article, the author can add a citation.
 If an Inbox is discovered in the cited article, dokieli sends a notification there to indicate what part of the article was cited by whom and where.
-dokieli-enabled articles also consume citation notifications to display these metrics for the author and other readers (\empty figure 4).
+dokieli-enabled articles also consume citation notifications to display these metrics for the author and other readers (figure~\ref{fig:like}).
 
-                                \empty Video of dokieli Web Annotation
-
-                                \empty Video of dokieli Share
+\begin{figure}
+  \begin{minipage}[b]{.49\textwidth}
+    \includegraphics[width=\textwidth]{media/images/dokieli-annotation}
+    \caption{Video of dokieli Web Annotation}
+    \label{fig:video-annotation}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{.49\textwidth}
+    \includegraphics[width=\textwidth]{media/images/dokieli-share}
+    \caption{Video of dokieli Share}
+    \label{fig:video-share}
+  \end{minipage}
+\end{figure}
 
@@
   \item If a {\em Like} activity is generated on the server, sloph uses the {\tt object} of the {\em Like} as the target for a notification.
-Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0 [\empty 18]), if the target is a dokieli article, this {\em Like} will be displayed (\empty figure 5).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e.
+Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0~\cite{ref-18}), if the target is a dokieli article, this {\em Like} will be displayed (figure~\ref{fig:like}).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e.
 places behind access control) any that it cannot automatically verify refer to third-party content published on another domain.
@@
 
-                                        Comparison of notification mechanisms
-
-                                                Mechanism
-                                                T1
-                                                T2
-                                                T3
-                                                R1
-                                                R2
-                                                R3
-                                                R4-A
-                                                R4-B
-                                                R4-Cp
-                                                R4-Cv
-                                                R4-Co
-                                                R5
-
-                                                    \begin{description}
-  \item[T1] Notification type\item[T2] Delivery method\item[T3] Dependencies\item[R1] Modularity (application classes: S Sender, R Receiver, C Consumer, U Subscriber/User)\item[R2] Reusability\item[R3] Persistence - required?
-how?\item[R4-A] Target representation\item[R4-B] Notification body\item[R4-Cp] Payload processing required?\item[R4-Cv] Verification - required?
-how?\item[R4-Co] Requirements for referenced resources?\item[R5] Subscription
-    \end{description}
-
-                                                    \begin{description}
-  \item[–] not applicable, out of scope\item[/] not specified, in scope\item[X] explicitly disallowed\item[app] application specific decision\item[!] required ({\em MUST})\item[+] recommended ({\em SHOULD})\item[O] optional ({\em MAY})\item[PuSH] PubSubHubbub
-    \end{description}
-
-                                                    \begin{description}
-  \item[h] HTML recommended\item[j] Alternate RDF formats can be negotiated\item[k] {\tt source} and {\tt target} key–value pairs is required\item[q] Provenance records with \empty PROV Ontology\item[r] RDF representation recommended\item[ra] SPARQL results transformed to RSS/Atom\item[s] \empty Sitemaps\item[t] Described in an RDF store or dataset
-    \end{description}
-
-                                                Semantic Pingback
-                                                Linkback
-                                                POST
-                                                RDF
-                                                S R
-                                                /
-                                                /
-                                                Anyr
-                                                form urlencodedk
-                                                !
-                                                ! parse source
-                                                Anyr
-                                                X
-
-                                                Webmention
-                                                Linkback
-                                                POST
-                                                HTML
-                                                S R
-                                                –
-                                                –
-                                                Anyh
-                                                form urlencodedk
-                                                !
-                                                ! parse source
-                                                Anyh
-                                                X
-
-                                                Provenance Pingback
-                                                Linkback
-                                                POST
-                                                RDF
-                                                S R
-                                                /
-                                                /
-                                                /
-                                                URI list
-                                                /
-                                                /
-                                                RDFq
-                                                X
-
-                                                DSNotify
-                                                Fat ping
-                                                POST, PUT
-                                                XML, PuSH
-                                                S U
-                                                /
-                                                –
-                                                –
-                                                XML
-                                                /
-                                                –
-                                                RDFt
-                                                !
-
-                                                sparqlPuSH
-                                                Fat ping
-                                                POST
-                                                XML, SPARQL, PuSH
-                                                S U
-                                                –
-                                                –
-                                                –
-                                                XMLra
-                                                /
-                                                –
-                                                RDFt
-                                                !
-
-                                                ResourceSync
-                                                Fat ping
-                                                POST
-                                                XML, PuSH
-                                                S U
-                                                /
-                                                –
-                                                –
-                                                XMLs
-                                                /
-                                                –
-                                                ?
-                                                !
-
-                                                Linked Data Notifications
-                                                Fat ping
-                                                POST
-                                                JSON-LD
-                                                S R C
-                                                !
-                                                ! URI
-                                                Any
-                                                JSON-LDj
-                                                + app
-                                                + app
-                                                –
-                                                O app
+\begin{table}
+  \caption{Comparison of notification mechanisms}
+  \label{tab:comparison}
+  \footnotesize
+  \begin{tabular}{% Mechanism
+    p{.14\textwidth}
+    % T1–T3
+    p{.065\textwidth}
+    p{.075\textwidth}
+    p{.1\textwidth}
+    % R1–R3
+    p{.03\textwidth}
+    p{.03\textwidth}
+    p{.04\textwidth}
+    % R4-A–R5
+    p{.06\textwidth}
+    p{.14\textwidth}
+    p{.045\textwidth}
+    p{.06\textwidth}
+    p{.06\textwidth}
+    p{.05\textwidth}
+    }\FL
+                                                Mechanism&
+                                                T1&
+                                                T2&
+                                                T3&
+                                                R1&
+                                                R2&
+                                                R3&
+                                                R4-A&
+                                                R4-B&
+                                                R4-C\textsuperscript{p}&
+                                                R4-C\textsuperscript{v}&
+                                                R4-C\textsuperscript{o}&
+                                                R5\ML
+                                                Semantic\newline Pingback&
+                                                Link-\newline back&
+                                                POST&
+                                                RDF&
+                                                S\newline R&
+                                                /&
+                                                /&
+                                                Any\textsuperscript{r}&
+                                                form\newline urlencoded\textsuperscript{k}&
+                                                !&
+                                                !\newline parse\newline source&
+                                                Any\textsuperscript{r}&
+                                                X\NN
+                                                Web-\newline mention&
+                                                Link-\newline back&
+                                                POST&
+                                                HTML&
+                                                S\newline R&
+                                                –&
+                                                –&
+                                                Any\textsuperscript{h}&
+                                                form\newline urlencoded\textsuperscript{k}&
+                                                !&
+                                                !\newline parse\newline source&
+                                                Any\textsuperscript{h}&
+                                                X\NN
+                                                Provenance\newline Pingback&
+                                                Link-\newline back&
+                                                POST&
+                                                RDF&
+                                                S\newline R&
+                                                /&
+                                                /&
+                                                /&
+                                                URI list&
+                                                /&
+                                                /&
+                                                RDF\textsuperscript{q}&
+                                                X\NN
+                                                DSNotify&
+                                                Fat\newline ping&
+                                                POST,\newline PUT&
+                                                XML,\newline PuSH&
+                                                S\newline U&
+                                                /&
+                                                –&
+                                                –&
+                                                XML&
+                                                /&
+                                                –&
+                                                RDF\textsuperscript{t}&
+                                                !\NN
+                                                sparqlPuSH&
+                                                Fat\newline ping&
+                                                POST&
+                                                XML,\newline SPARQL,\newline PuSH&
+                                                S\newline U&
+                                                –&
+                                                –&
+                                                –&
+                                                XML\textsuperscript{ra}&
+                                                /&
+                                                –&
+                                                RDF\textsuperscript{t}&
+                                                !\NN
+                                                Resource-\newline Sync&
+                                                Fat\newline ping&
+                                                POST&
+                                                XML,\newline PuSH&
+                                                S\newline U&
+                                                /&
+                                                –&
+                                                –&
+                                                XML\textsuperscript{s}&
+                                                /&
+                                                –&
+                                                ?&
+                                                !\NN
+                                                Linked Data\newline Notifications&
+                                                Fat\newline ping&
+                                                POST&
+                                                JSON-LD&
+                                                S\newline R\newline C&
+                                                !&
+                                                !\newline URI&
+                                                Any&
+                                                JSON-LD\textsuperscript{j}&
+                                                + app&
+                                                + app&
+                                                –&
+                                                O\newline app\LL
+  \end{tabular}\\
+  \begin{tabularx}{1.0\linewidth}{XX}\FL
+    \textbf{T1}: Notification type&
+    \textbf{R4-A}: Target representation\NN
+    \textbf{T2}: Delivery method&
+    \textbf{R4-B}: Notification body\NN
+    \textbf{T3}: Dependencies&
+    \textbf{R4-C\textsuperscript{p}}: Payload processing required?\NN
+    \textbf{R1}: Modularity (application classes: S Sender, R Receiver, C Consumer, U Subscriber/User)&
+    \textbf{R4-C\textsuperscript{v}}: Verification - required? how?\NN
+    \textbf{R2}: Reusability&
+    \textbf{R4-C\textsuperscript{o}}: Requirements for referenced resources?\NN
+    \textbf{R3}: Persistence - required? how?&
+    \textbf{R5}: Subscription\ML
+    \textbf{–}: not applicable, out of scope&
+    \textbf{!}: required ({\em MUST})\NN
+    \textbf{/}: not specified, in scope&
+    \textbf{+}: recommended ({\em SHOULD})\NN
+    \textbf{X}: explicitly disallowed&
+    \textbf{O}: optional ({\em MAY})\NN
+    \textbf{app}: application specific decision&
+    \textbf{PuSH}: PubSubHubbub\ML
+    \textbf{\textsuperscript{h}}: HTML recommended&
+    \textbf{\textsuperscript{r}}: RDF representation recommended\NN
+    \textbf{\textsuperscript{j}}: Alternate RDF formats can be negotiated&
+    \textbf{\textsuperscript{ra}}: SPARQL results transformed to RSS/Atom\NN
+    \textbf{\textsuperscript{k}}: {\tt source} and {\tt target} key–value pairs is required&
+    \textbf{\textsuperscript{s}}: \empty Sitemaps\NN
+    \textbf{\textsuperscript{q}}: Provenance records with \empty PROV Ontology&
+    \textbf{\textsuperscript{t}}: Described in an RDF store or dataset\LL
+  \end{tabularx}
+\end{table}
 
@@
 
-                            \begin{enumerate}
-  \item Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016, \empty http://www2016.net/proceedings/companion/p223.pdf\item Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012, \empty \item Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016, \empty http://events.linkeddata.org/ldow2016/papers/LDOW2016\_paper\_02.pdf\item Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/ldn/\item Parecki, A.: Webmention, W3C Proposed Recommendation, 2016, \empty https://www.w3.org/TR/webmention/\item Langridge, S., Hickson, I.: Pingback 1.0, 2002, \empty http://www.hixie.ch/specs/pingback/pingback\item Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \empty http://www.w3.org/TR/prov-aq/\item Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010, \empty http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf\item Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol.
-699, 2010, \empty http://ceur-ws.org/Vol-699/Paper6.pdf\item Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016, \empty http://www.openarchives.org/rs/notification/1.0/notification\item Berners-Lee, T.: Cool URIs don't change, W3C, 1998, \empty https://www.w3.org/Provider/Style/URI.html\item Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On Persistent URIs, 2012, \empty http://philarcher.org/diary/2013/uripersistence/\item Fielding, R.
+\paragraph{Acknowledgements:}
+This material is based on work supported by the Qatar Computing Research Institute (QCRI), and by the DFG project “Opening Scholarly Communication in Social Sciences” (grant agreement AU 340/9-1).
+
+\begin{thebibliography}{99}
+  \bibitem{ref-1} Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016%, \url{http://www2016.net/proceedings/companion/p223.pdf}
+  \bibitem{ref-2} Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012
+  \bibitem{ref-3} Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016%, \url{http://events.linkeddata.org/ldow2016/papers/LDOW2016_paper_02.pdf}
+  \bibitem{ref-4} Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/ldn/}
+  \bibitem{ref-5} Parecki, A.: Webmention, W3C Proposed Recommendation, 2016,  \url{https://www.w3.org/TR/webmention/}
+  \bibitem{ref-6} Langridge, S., Hickson, I.: Pingback 1.0, 2002,  \url{http://www.hixie.ch/specs/pingback/pingback}
+  \bibitem{ref-7} Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \url{http://www.w3.org/TR/prov-aq/}
+  \bibitem{ref-8} Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010%, \url{http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf}
+  \bibitem{ref-9} Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol.
+699, 2010%, \url{http://ceur-ws.org/Vol-699/Paper6.pdf}
+  \bibitem{ref-10} Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016,  \url{http://www.openarchives.org/rs/notification/1.0/notification}
+  \bibitem{ref-11} Berners-Lee, T.: Cool URIs don't change, W3C, 1998,  \url{https://www.w3.org/Provider/Style/URI.html}
+  \bibitem{ref-12} Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On Persistent URIs, 2012, \url{http://philarcher.org/diary/2013/uripersistence/}
+  \bibitem{ref-13} Fielding, R.
 T.: Architectural Styles and the Design of Network-based Software Architectures.
-Doctoral dissertation, University of California, Irvine, 2000, \empty http://www.ics.uci.edu/\~fielding/pubs/dissertation/rest\_arch\_style.htm\item Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011, \empty http://ws-rest.org/2011/proc/a5-page.pdf\item Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016, \empty \item Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C Recommendation, 2015, \empty https://www.w3.org/TR/ldp/\item Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016, \empty http://csarven.ca/dokieli\item Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitystreams-core/\item Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitypub/\item Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014, \empty http://manu.sporny.org/2014/json-ld-origins-2/
-    \end{enumerate}
+Doctoral dissertation, University of California, Irvine, 2000%,  \url{http://www.ics.uci.edu/~fielding/pubs/dissertation/rest_arch_style.htm}
+  \bibitem{ref-14} Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011%,  \url{http://ws-rest.org/2011/proc/a5-page.pdf}
+  \bibitem{ref-15} Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016
+  \bibitem{ref-16} Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C Recommendation, 2015, \url{https://www.w3.org/TR/ldp/}
+  \bibitem{ref-17} Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016, \url{http://csarven.ca/dokieli}
+  \bibitem{ref-18} Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/activitystreams-core/}
+  \bibitem{ref-19} Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/activitypub/}
+  \bibitem{ref-20} Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014, \url{http://manu.sporny.org/2014/json-ld-origins-2/}
+\end{thebibliography}
 

### 2017-03-17 index.tex
minor shortening back to 16 pages

@@
 
-\par LDN is a W3C \empty Candidate Recommendation via the \empty Social Web Working Group~\cite{ref-4}. The first two authors of this article are the co-editors of the specification.
+\par LDN is a W3C \empty Candidate Recommendation via the \empty Social Web Working Group~\cite{ref-4}. The first two authors of this article co-edited the specification.
 
@@
   \bibitem{ref-8} Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010%, \url{http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf}
-  \bibitem{ref-9} Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol.
+  \bibitem{ref-9} Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR-WS.org, Vol.
 699, 2010%, \url{http://ceur-ws.org/Vol-699/Paper6.pdf}

### 2017-03-22 index.tex
Mention the initial work on the Inbox in acknowledgements

@@
 \paragraph{Acknowledgements:}
+The initial work on the Inbox was done by the Solid project.
 This material is based on work supported by the Qatar Computing Research Institute (QCRI), and by the DFG project “Opening Scholarly Communication in Social Sciences” (grant agreement AU 340/9-1).
